### 2018-06-06 03-Details.Rmd
adjusted travis, fixed bug in report

# Details

Suppose we want to solve the following optimization problem:

\begin{align*}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{align*}

where $x \in \mathbb{R}^{n}, z \in \mathbb{R}^{m}, A \in \mathbb{R}^{p \times n}, B \in \mathbb{R}^{p \times m}$, and $c \in \mathbb{R}^{p}$. Following @boyd2011distributed, the optimization problem will be introduced in vector form though we will later consider cases where $x$ and $z$ are matrices. We will also assume $f$ and $g$ are convex functions. Optimization problems like this arise naturally in several statistics and machine learning applications -- particularly regularization methods. For instance, we could take $f$ to be the squared error loss, $g$ to be the $l_{2}$-norm, $c$ to be equal to zero and $A$ and $B$ to be identity matrices to solve the ridge regression optimization problem.

The *augmented lagrangian* is constructed as follows:

\[ L_{\rho}(x, z, y) = f(x) + g(z) + y^{T}(Ax + Bz - c) + \frac{\rho}{2}\left\| Ax + Bz - c \right\|_{2}^{2} \]

where $y \in \mathbb{R}^{p}$ is the lagrange multiplier and $\rho > 0$ is a scalar. Clearly any minimizer, $p^{*}$, under the augmented lagrangian is equivalent to that of the lagrangian since any feasible point $(x, z)$ satisfies the constraint $\rho\left\| Ax + Bz - c \right\|_{2}^{2}/2 = 0$.

\[ p^{*} = inf\left\{ f(x) + g(z) | Ax + Bz = c \right\} \]

The alternating direction method of multipliers (ADMM) algorithm consists of the following repeated iterations:

\begin{align}
  x^{k + 1} &:= \arg\min_{x}L_{\rho}(x, z^{k}, y^{k}) \\
  z^{k + 1} &:= \arg\min_{z}L_{\rho}(x^{k + 1}, z, y^{k}) \\
  y^{k + 1} &:= y^{k} + \rho(Ax^{k + 1} + Bz^{k + 1} - c)
\end{align}

A more complete introduction to the algorithm -- specifically how it arose out of *dual ascent* and *method of multipliers* -- can be found in @boyd2011distributed.

<br>\vspace{1cm}

## ADMM Algorithm

Now consider the case where $X_{1}, ..., X_{n}$ are iid $N_{p}(\mu, \Sigma)$ random variables and we are tasked with estimating the precision matrix, denoted $\Omega \equiv \Sigma^{-1}$. The maximum likelihood estimator for $\Omega$ is

\[ \hat{\Omega}_{MLE} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ Tr\left(S\Omega\right) - \log \det\left(\Omega \right) \right\} \]

where $S = \sum_{i = 1}^{n}(X_{i} - \bar{X})(X_{i} - \bar{X})^{T}/n$ and $\bar{X}$ is the sample mean. By setting the gradient equal to zero, we can show that when the solution exists, $\hat{\Omega}_{MLE} = S^{-1}$.

As in regression settings, we can construct a *penalized* likelihood estimator by adding a penalty term, $P\left( \Omega \right)$, to the likelihood:

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ Tr\left(S\Omega\right) - \log \det\left(\Omega \right) + P\left( \Omega \right) \right\} \]

$P\left( \Omega \right)$ is often of the form $P\left(\Omega \right) = \lambda\|\Omega \|_{F}^{2}$ or $P\left(\Omega \right) = \|\Omega\|_{1}$ where $\lambda > 0$, $\left\|\cdot \right\|_{F}^{2}$ is the Frobenius norm and we define $\left\|A \right\|_{1} = \sum_{i, j} \left| A_{ij} \right|$. These penalties are the ridge and lasso, respectively. We will, instead, take $P\left( \Omega \right) = \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right]$ so that the full penalized likelihood is

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ Tr\left(S\Omega\right) - \log \det\left(\Omega \right) + \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \right\} \]

where $0 \leq \alpha \leq 1$. This *elastic-net* penalty was explored by Hui Zou and Trevor Hastie [@zou2005regularization] and is identical to the penalty used in the popular penalized regression package `glmnet`. Clearly, when $\alpha = 0$ the elastic-net reduces to a ridge-type penalty and when $\alpha = 1$ it reduces to a lasso-type penalty.

By letting $f$ be equal to the non-penalized likelihood and $g$ equal to $P\left( \Omega \right)$, our goal is to minimize the full augmented lagrangian where the constraint is that $\Omega - Z$ is equal to zero:

\[ L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + Tr\left[\Lambda\left(\Omega - Z\right)\right] + \frac{\rho}{2}\left\|\Omega - Z\right\|_{F}^{2} \]

The ADMM algorithm for estimating the penalized precision matrix in this problem is

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega}\left\{ Tr\left(S\Omega\right) - \log\det\left(\Omega\right) + Tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + Tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
\end{align}

<br>\vspace{1cm}

## Scaled-Form ADMM

An alternate form of the ADMM algorithm can constructed by scaling the dual variable ($\Lambda^{k}$). Let us define $R^{k} = \Omega - Z^{k}$ and $U^{k} = \Lambda^{k}/\rho$.

\begin{align*}
  Tr\left[ \Lambda^{k}\left( \Omega - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} &= Tr\left[ \Lambda^{k}R^{k} \right] + \frac{\rho}{2}\left\| R^{k} \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + \Lambda^{k}/\rho \right\|_{F}^{2} - \frac{\rho}{2}\left\| \Lambda^{k}/\rho \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + U^{k} \right\|_{F}^{2} - \frac{\rho}{2}\left\| U^{k} \right\|_{F}^{2}
\end{align*}

Therefore, the condensed-form ADMM algorithm can now be written as

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega}\left\{ Tr\left(S\Omega\right) - \log\det\left(\Omega\right) + \frac{\rho}{2}\left\| \Omega - Z^{k} + U^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z + U^{k} \right\|_{F}^{2} \right\} \\
  U^{k + 1} &= U^{k} + \Omega^{k + 1} - Z^{k + 1}
\end{align}

And more generally (in vector form),

\begin{align}
  x^{k + 1} &= \arg\min_{x}\left\{ f(x) + \frac{\rho}{2}\left\| Ax + Bz^{k} - c + u^{k} \right\|_{2}^{2} \right\} \\
  z^{k + 1} &= \arg\min_{z}\left\{ g(z) + \frac{\rho}{2}\left\| Ax^{k + 1} + Bz - c + u^{k} \right\|_{2}^{2} \right\} \\
  u^{k + 1} &= u^{k} + Ax^{k + 1} + Bz^{k + 1} - c
\end{align}

Note that there are limitations to using this method. Because the dual variable is scaled by $\rho$ (the step size), this form limits one to using a constant step size without making further adjustments to $U^{k}$. It has been shown in the literature that a dynamic step size can significantly reduce the number of iterations required for convergence.

<br>\vspace{1cm}

## Algorithm

\begin{align*}
  \Omega^{k + 1} &= \arg\min_{\Omega}\left\{ Tr\left(S\Omega\right) - \log\det\left(\Omega\right) + Tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + Tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
\end{align*}

<br>\vspace{1cm}

Initialize $Z^{0}, \Lambda^{0}$, and $\rho$. Iterate the following three steps until convergence:

1. Decompose $S + \Lambda^{k} - \rho Z^{k} = VQV^{T}$ via spectral decomposition.

\[ \Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + \left( Q^{2} + 4\rho I_{p} \right)^{1/2} \right]V^{T} \]

2. Elementwise soft-thresholding for all $i = 1,..., p$ and $j = 1,..., p$.

\begin{align*}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}Sign\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}Soft\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
\end{align*}

3. Update $\Lambda^{k + 1}$.

\[ \Lambda^{k + 1} = \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right) \]

<br>\vspace{1cm}

### Proof of (1):

\[ \Omega^{k + 1} = \arg\min_{\Omega}\left\{ Tr\left(S\Omega\right) - \log\det\left(\Omega\right) + Tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \]

\begin{align*}
  \nabla_{\Omega}&\left\{ Tr\left(S\Omega\right) - \log\det\left(\Omega\right) + Tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  &= S - \Omega^{-1} + \Lambda^{k} + \rho\left( \Omega - Z^{k} \right)
\end{align*}

Set the gradient equal to zero and decompose $\Omega = VDV^{T}$ where $D$ is a diagonal matrix with diagonal elements equal to the eigen values of $\Omega$ and $V$ is the matrix with corresponding eigen vectors as columns.

\[ S + \Lambda^{k} - \rho Z^{k} = \Omega^{-1} - \rho \Omega = VD^{-1}V^{T} - \rho VDV^{T} =  V\left(D^{-1} - \rho D\right)V^{T} \]

This equivalence implies that

\[ \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right) = \frac{1}{\phi_{j}(\Omega^{k + 1})} - \rho\phi_{j}(\Omega^{k + 1}) \]

where $\phi_{j}(\cdot)$ is the $j$th eigen value.

\begin{align*}
  &\Rightarrow \rho\phi_{j}^{2}(\Omega^{k + 1}) + \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right)\phi_{j}(\Omega^{k + 1}) - 1 = 0 \\
  &\Rightarrow \phi_{j}(\Omega^{k + 1}) = \frac{-\phi_{j}(S + \Lambda^{k} - \rho Z^{k}) \pm \sqrt{\phi_{j}^{2}(S + \Lambda^{k} - \rho Z^{k}) + 4\rho}}{2\rho}
\end{align*}

In summary, if we decompose $S + \Lambda^{k} - \rho Z^{k} = VQV^{T}$ then

\[ \Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + (Q^{2} + 4\rho I_{p})^{1/2}\right] V^{T} \]

<br>\vspace{1cm}

### Proof of (2)

\[ Z^{k + 1} = \arg\min_{Z}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + Tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \]

\begin{align*}
  \partial&\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + Tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \partial\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + Tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right]\right\} + \nabla_{\Omega}\left\{\frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \lambda(1 - \alpha)Z + Sign(Z)\lambda\alpha - \Lambda^{k} - \rho\left( \Omega^{k + 1} - Z \right)
\end{align*}

where $Sign(Z)$ is the elementwise Sign operator. By setting the gradient/sub-differential equal to zero, we arrive at the following equivalence:

\[ Z_{ij} = \frac{1}{\lambda(1 - \alpha) + \rho}\left( \rho \Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} - Sign(Z_{ij})\lambda\alpha \right) \]

for all $i = 1,..., p$ and $j = 1,..., p$. We observe two scenarios:

- If $Z_{ij} > 0$ then

\[ \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} > \lambda\alpha \]

- If $Z_{ij} < 0$ then

\[ \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} < -\lambda\alpha \]

This implies that $Sign(Z_{ij}) = Sign(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k})$. Putting all the pieces together, we arrive at

\begin{align*}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}Sign\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}Soft\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
\end{align*}

where $Soft$ is the soft-thresholding function.

<br><br>\newpage

<!-- \begin{thebibliography}{9} -->

<!-- \bibitem{1} -->
<!--   Boyd, Stephen, et al. "Distributed optimization and statistical learning via the alternating direction method of multipliers." Foundations and Trends® in Machine Learning 3.1 (2011): 1-122. -->

<!-- \bibitem{2} -->
<!--   Polson, Nicholas G., James G. Scott, and Brandon T. Willard. "Proximal algorithms in statistics and machine learning." Statistical Science 30.4 (2015): 559-581. -->

<!-- \bibitem{3} -->
<!--   Marjanovic, Goran, and Victor Solo. "On $ l_q $ optimization and matrix completion." IEEE Transactions on signal processing 60.11 (2012): 5714-5724. -->

<!-- \bibitem{4} -->
<!--   Zou, Hui, and Trevor Hastie. "Regularization and variable selection via the elastic net." Journal of the Royal Statistical Society: Series B (Statistical Methodology) 67.2 (2005): 301-320. -->

<!-- \end{thebibliography} -->

## References