### 2018-06-06 02-Tutorial.Rmd
Deploy MGallow/oral_manuscript to github.com/MGallow/oral_manuscript.git:master

# Tutorial

<br>\vspace{0.5cm}
```{r, eval = FALSE, echo = TRUE}
# The easiest way to install is from CRAN
install.packages("ADMMsigma")

# You can also install the development version from GitHub:
# install.packages("devtools")
devtools::install_github("MGallow/ADMMsigma")
```
<br>\vspace{0.5cm}

If there are any issues/bugs, please let me know: [github](https://github.com/MGallow/ADMMsigma/issues). You can also contact me via my [website](https://mgallow.github.io/). Pull requests are welcome!

<br>\vspace{0.5cm}

A (possibly incomplete) list of functions contained in the package can be found below:

* `ADMMsigma()` computes the estimated precision matrix (ridge, lasso, and elastic-net type regularization optional)

* `RIDGEsigma()` computes the estimated ridge penalized precision matrix via closed-form solution

* `plot.ADMMsigma()` produces a heat map or line graph for cross validation errors

* `plot.RIDGEsigma()` produces a heat map or line graph for cross validation errors

<br>\vspace{1cm}

## Usage

We will first generate data from a sparse, tri-diagonal precision matrix and denote it as Omega.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}
library(ADMMsigma)

#  generate data from a sparse matrix
# first compute covariance matrix
S = matrix(0.7, nrow = 5, ncol = 5)
for (i in 1:5){
  for (j in 1:5){
    S[i, j] = S[i, j]^abs(i - j)
  }
}

# print oracle precision matrix (shrinkage might be useful)
(Omega = round(qr.solve(S), 3))

# generate 100 x 5 matrix with rows drawn from iid N_p(0, S)
set.seed(123)
Z = matrix(rnorm(100*5), nrow = 100, ncol = 5)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

# snap shot of data
head(X)

```
<br>\vspace{0.5cm}

As described earlier in the report, the maximum likelihood estimator (MLE) for Omega is the inverse of the sample precision matrix $S^{-1} = \left[\sum_{i = 1}^{n}(X_{i} - \bar{X})(X_{i} - \bar{X})^{T}/n \right]^{-1}$:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# print inverse of sample precision matrix (perhaps a bad estimate)
round(qr.solve(cov(X)*(nrow(X) - 1)/nrow(X)), 5)

```
<br>\vspace{0.5cm}

However, because Omega (known as the *oracle*) is sparse, a shrinkage estimator will perhaps perform better than the sample estimator. Below we construct various penalized estimators:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# elastic-net type penalty (set tolerance to 1e-8)
ADMMsigma(X, tol.abs = 1e-8, tol.rel = 1e-8)

```
<br>\vspace{0.5cm}

**LASSO:**

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# lasso penalty (default tolerance)
ADMMsigma(X, alpha = 1)

```
<br>\vspace{0.5cm}

**ELASTIC-NET:**

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# elastic-net penalty (alpha = 0.5)
ADMMsigma(X, alpha = 0.5)

```
<br>\newpage

**RIDGE:**

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# ridge penalty
ADMMsigma(X, alpha = 0)

# ridge penalty no ADMM
RIDGEsigma(X, lam = 10^seq(-8, 8, 0.01))

```
<br>\newpage

This package also has the capability to provide heat maps for the cross validation errors. The more bright (white) areas of the heat map pertain to more optimal tuning parameters.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# produce CV heat map for ADMMsigma
ADMM = ADMMsigma(X, tol.abs = 1e-8, tol.rel = 1e-8)
plot(ADMM, type = "heatmap")

```
<br>\newpage

```{r, message = FALSE, echo = TRUE}

# produce line graph for CV errors for ADMMsigma
plot(ADMM, type = "line")

```
<br>\newpage

```{r, message = FALSE, echo = TRUE}

# produce CV heat map for RIDGEsigma
RIDGE = RIDGEsigma(X, lam = 10^seq(-8, 8, 0.01))
plot(RIDGE, type = "heatmap")

```

<br>\newpage
```{r, message = FALSE, echo = TRUE}

# produce line graph for CV errors for RIDGEsigma
plot(RIDGE, type = "line")

```