### 2018-07-17 06-SCPME-Details.Rmd

# `SCPME` Details

Consider the case where we observe $n$ independent, identically distributed copies of the random variable ($X_{i}$) where $X_{i} \in \mathbb{R}^{p}$ is normally distributed with some mean, $\mu$, and some variance, $\Sigma$. That is, $X_{i} \sim N_{p}\left( \mu, \Sigma \right)$.

Because we assume independence, we know that the probability of observing these specific observations $X_{1}, ..., X_{n}$ is equal to

\begin{align*}
  f(X | \mu, \Sigma) &= \prod_{i = 1}^{n}(2\pi)^{-p/2}\left| \Sigma \right|^{-1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)^{T}\Sigma^{-1}\left( X_{i} - \mu \right) \right] \\
  &= (2\pi)^{-nr/2}\left| \Sigma \right|^{-n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu \right)\left( X_{i} - \mu \right)^{T}\Sigma^{-1} \right]
\end{align*}

where $\mbox{etr}\left( \cdot \right)$ denotes the exponential trace operator. It follows that the log-likelihood for $\mu$ and $\Sigma$ is equal to the following:

\[ l(\mu, \Sigma | X) = const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-1} \right] \]

If we are interested in estimating $\mu$, it is relatively straight forward to show that the maximum likelihood estimator (MLE) for $\mu$ is $\hat{\mu} = \sum_{i = 1}^{n}X_{i}/n \equiv \bar{X}$. However, in addition to $\mu$, many practical applications require the estimation of $\Sigma$ as well. Like $\mu$, we can find a maximum likelihood estimator:

\begin{align*}
  &\hat{\Sigma} = \arg\max_{\Sigma \in \mathbb{S}_{+}^{p}}\left\{ const. - \frac{n}{2}\log\left| \Sigma \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-1} \right] \right\} \\
  &\nabla_{\Sigma}l(\mu, \Sigma | X) = -\frac{n}{2}\Sigma^{-1} - \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)^{T}\Sigma^{-2} \\
  \Rightarrow &\hat{\Sigma} = \left[ \frac{1}{n}\sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)^{T} \right]
\end{align*}

by setting the gradient equal to zero and plugging in the MLE for $\mu$. This implies that the MLE for $\Sigma$ is our usual sample estimator often denoted as $S$. We could have just as easily computed the maximum likelihood estimator for the precision matrix $\Omega \equiv \Sigma^{-1}$:

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| \right\} \]

so that $\hat{\Omega} = S^{-1}$. Notice that here we are *minimizing* the negative log-likelihood as opposed to maximizing the log-likelihood. Both procedures will result in the same estimate. Beyond the formatting convenience, computing estimates for $\Omega$ as opposed to $\Sigma$ often poses less computational challenges -- and accordingly, the literature has placed more emphasis on efficiently solving for $\Omega$ instad of $\Sigma$.

As in regression settings, we can construct a *penalized* log-likelihood estimator by adding a penalty term, $P\left(\Omega\right)$, to the likelihood:

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\} \]

$P\left( \Omega \right)$ is often of the form $P\left(\Omega \right) = \lambda\|\Omega \|_{F}^{2}$ or $P\left(\Omega \right) = \|\Omega\|_{1}$ where $\lambda > 0$, $\left\|\cdot \right\|_{F}^{2}$ is the Frobenius norm and we define $\left\|A \right\|_{1} = \sum_{i, j} \left| A_{ij} \right|$. These penalties are the ridge and lasso, respectively. The penalty proposed by @molstad2017shrinking is one of the following form:

\[ P\left(\Omega\right) = \lambda\left\| A\Omega B - C \right\|_{1} \]

so that solving the full penalized log-likelihood for $\Omega$ results in solving

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left\| A\Omega B - C \right\|_{1} \right\} \]

where $A \in \mathbb{R}^{n \times p}, B \in \mathbb{R}^{p \times r}, \mbox{ and } C \in \mathbb{R}^{n \times r}$ are matrices assumed to be known and specified by the user. As discussed [here](https://mgallow.github.io/SCPME/articles/Jointloglik.html), this type of penalty leads to many new, interesting, and novel estimators for $\Omega$. An example of one such estimator would be one where we set $A = I_{p}, B = \Sigma_{xy}, \mbox{ and } C = 0$ so that we assume sparsity in the forward regression coefficient $\beta \equiv \Omega\Sigma_{xy}$.

We will explore how to solve for $\hat{\Omega}$ in the next section.

<br>\vspace{1cm}

## Augmented ADMM Algorithm

This section requires general knowledge of the ADMM algorithm. I would recommend reading this overview I have written [here](https://mgallow.github.io/ADMMsigma/articles/Details.html#admm-algorithm) before proceeding.

The ADMM algorithm within the context of this problem consists of the following repeated iterations:

\begin{align}
  \Omega^{k + 1} &:= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}L_{\rho}(\Omega, Z^{k}, \Lambda^{k}) \\
  Z^{k + 1} &:= \arg\min_{Z \in \mathbb{R}^{n \times r}}L_{\rho}(\Omega^{k + 1}, Z, \Lambda^{k}) \\
  \Lambda^{k + 1} &:= \Lambda^{k} + \rho\left(A\Omega^{k + 1}B - Z^{k + 1} - C \right)
\end{align}

where $L_{p}(\cdot)$ is the *augmented lagrangian* defined as

\[ L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda^{T}\left(A\Omega B - Z - C\right)\right] + \frac{\rho}{2}\left\|A\Omega B - Z - C\right\|_{F}^{2} \]

with $f\left(\Omega\right) = tr\left(S\Omega\right) - \log\left|\Omega\right|$ and $g\left(Z\right) = \lambda\left\|Z\right\|_{1}$.

We instead replace with

\begin{align*}
  tr\left(S\Omega\right) &- \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C) \right] \\
  &+ \frac{\rho}{2}\left\|A\Omega B - Z^{k} - C \right\|_{F}^{2} + \frac{\rho}{2}vec\left(\Omega - \Omega^{k}\right)^{T}Q\left(\Omega - \Omega^{k}\right) \\
  &= tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})^{T}(A\Omega B - Z^{k} - C)\right] + \frac{\rho}{2}\left\|A\Omega B - Z^{k} - C\right\|_{F}^{2} \\
  &+ \frac{\rho}{2}tr\left[A^{T}A(\Omega - \Omega^{k})BB^{T}(\Omega - \Omega^{k})\right] + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k}\right\|_{F}^{2}
\end{align*}

**Note:**

\begin{align*}
  vec\left(\Omega - \Omega^{k}\right)^{T}\left(A^{T}A \otimes BB^{T}\right)vec\left(\Omega - \Omega^{k}\right) &= vec\left(\Omega - \Omega^{k}\right)^{T}vec\left(BB^{T}(\Omega - \Omega^{k})A^{T}A\right) \\
  &= tr\left[A^{T}A\left(\Omega - \Omega^{k}\right)BB^{T}\left(\Omega - \Omega^{k}\right)\right]
\end{align*}

Suppose we define $G^{k} := A\Omega^{k}B - Z^{k} - C + \Lambda^{k}/\rho$. Then

\begin{align*}
  &\nabla_{\Omega}\left\{tr\left[\left(S + \rho A^{T}G^{k}B^{T}\right)\Omega\right] - \log\left|\Omega\right| + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k}\right\|_{F}^{2} \right\} \\
  &= 2S - S\circ I_{p} + G^{k} + (G^{k})^{T} - G^{k}\circ I_{p} - 2\Omega^{-1} + \Omega^{-1}\circ I_{p} \\
  &+ \frac{\rho\tau}{2}\left[2\Omega - 2(\Omega^{k})^{T} + 2\Omega^{T} - 2\Omega^{k} - 2(\Omega - \Omega^{k})^{T}\circ I_{p} \right]
\end{align*}

Note that we need to honor the symmetric constraint given by $\Omega$. By setting the gradient equal to zero and multiplying all off-diagonal elements by $1/2$, this simplifies to

\begin{align*}
  0 &= S + \frac{1}{2}\left(G^{k} + (G^{k})^{T}\right) - (\Omega^{k + 1})^{-1} + \rho\tau\left(\Omega^{k + 1} - \Omega^{k}\right) \\
  \Rightarrow &S + \frac{1}{2}\left(G^{k} + (G^{k})^{T}\right) - \rho\tau\Omega^{k} = (\Omega^{k + 1})^{-1} - \rho\tau\Omega^{k + 1}
\end{align*}

Consider

\begin{align*}
  &\nabla_{\Omega}\left\{ \frac{\rho}{2}\left\|A\Omega B - Z - C\right\|_{F}^{2} \right\} = \rho BB^{T}\Omega A^{T}A - \rho BZ^{T}A - \rho BC^{T}A \\
  &\nabla_{\Omega}^{2}\left\{ \frac{\rho}{2}\left\|A\Omega B - Z - C \right\|_{F}^{2} \right\} = \rho\left(A^{T}A \otimes BB^{T} \right)
\end{align*}

\begin{align*}
  \frac{\rho}{2}\left\| A\Omega B - Z^{k} - C \right\|_{F}^{2} &= \frac{\rho}{2}\left\| A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2} \\
  &+ \frac{\rho}{2}vec\left( \Omega - \Omega^{k}\right)^{T}\left(A^{T}A \otimes BB^{T}\right)vec\left(\Omega - \Omega^{k}\right) \\
  &+ \rho vec\left(\Omega - \Omega^{k}\right)^{T}vec\left(BB^{T}\Omega^{k}A^{T}A - B(Z^{k})^{T}A - BC^{T}A \right)
\end{align*}

Consider $\rho vec\left(\Omega - \Omega^{k}\right)^{T}Q vec\left(\Omega - \Omega^{k} \right)/2$ where $Q = \tau I_{p} - \left(A^{T}A \otimes BB^{T}\right)$ and $\tau$ is chosen such that $Q$ is positive definite.

Note that if $Q$ is positive definite (p.d.), then

\[ \frac{\rho}{2}vec\left(\Omega - \Omega^{k} \right)Q\left(\Omega - \Omega^{k} \right) > 0 \]

since $\rho > 0$ and $vec\left(\Omega - \Omega^{k}\right)$ is always nonzero whenever $\Omega \neq \Omega^{k}$.

This implies the following fact:

\[ \frac{\rho}{2}\left\|A\Omega B - Z^{k} - C \right\|_{F}^{2} \leq \frac{\rho}{2}\left\|A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2} + \frac{\rho}{2}vec\left(\Omega - \Omega^{k} \right)^{T}Q vec\left(\Omega - \Omega^{k} \right) \]

Further,

\begin{align*}
  \nabla_{\Omega}&\left\{const. + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k} \right\|_{F}^{2} + \rho tr\left[\Omega^{T}\left(BB^{T}\Omega^{k}A^{T}A - B(Z^{k})^{T}A - BC^{T}A \right) \right] \right\} \\
  &= \rho\tau\left(\Omega - \Omega^{k}\right) + \rho BB^{T}\Omega A^{T}A - \rho B(Z^{k})^{T}A - \rho BC^{T}A
\end{align*}

This implies a majorizing function!

<br>\vspace{1cm}

### Algorithm

Set $k = 0$ and repeat steps 1-6 until convergence.

1. Compute $G_{k} = \rho A^{T}\left( A\Omega_{k} B - Z_{k} - C + \rho^{-1}Y_{k} \right)$

2. Decompose $S + \left( G_{k} + G_{k}^{T} \right)/2 - \rho\tau\Omega_{k} = VQV^{T}$ (via the spectral decomposition).

3. Set $\Omega_{k + 1} = V\left( -Q + (Q^{2} + 4\rho\tau I_{p})^{1/2} \right)V^{T}/(2\rho\tau)$

4. Set $Z_{k + 1} = \mbox{soft}\left( A\Omega_{k + 1}B - C + \rho^{-1}Y_{k}, \rho^{-1}\lambda \right)$

5. Set $Y_{k + 1} = \rho\left( A\Omega_{k + 1} B - Z_{k + 1} - C \right)$

6. Replace $k$ with $k + 1$.

where $Soft(a, b) = Sign(a)(\left| a \right| - b)_{+}$.

<br>\vspace{1cm}

### Proof of (2-3):

\[ \Omega^{k + 1} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \]

\vspace{1cm}

\begin{align*}
  \nabla_{\Omega}&\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  &= S - \Omega^{-1} + \Lambda^{k} + \rho\left( \Omega - Z^{k} \right)
\end{align*}

Note that because all of the variables are symmetric, we can essentially ignore the constraint when deriving the gradient.
Set the gradient equal to zero and decompose $\Omega = VDV^{T}$ where $D$ is a diagonal matrix with diagonal elements equal to the eigen values of $\Omega$ and $V$ is the matrix with corresponding eigen vectors as columns.

\[ S + \Lambda^{k} - \rho Z^{k} = \Omega^{-1} - \rho \Omega = VD^{-1}V^{T} - \rho VDV^{T} =  V\left(D^{-1} - \rho D\right)V^{T} \]

This equivalence implies that

\[ \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right) = \frac{1}{\phi_{j}(\Omega^{k + 1})} - \rho\phi_{j}(\Omega^{k + 1}) \]

where $\phi_{j}(\cdot)$ is the $j$th eigen value.

\begin{align*}
  &\Rightarrow \rho\phi_{j}^{2}(\Omega^{k + 1}) + \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right)\phi_{j}(\Omega^{k + 1}) - 1 = 0 \\
  &\Rightarrow \phi_{j}(\Omega^{k + 1}) = \frac{-\phi_{j}(S + \Lambda^{k} - \rho Z^{k}) \pm \sqrt{\phi_{j}^{2}(S + \Lambda^{k} - \rho Z^{k}) + 4\rho}}{2\rho}
\end{align*}

In summary, if we decompose $S + \Lambda^{k} - \rho Z^{k} = VQV^{T}$ then

\[ \Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + (Q^{2} + 4\rho I_{p})^{1/2}\right] V^{T} \]

<br>\vspace{1cm}

### Proof of (4)

\[ Z^{k + 1} = \arg\min_{Z \in \mathbb{R}^{n \times r}}\left\{ \lambda\left\| Z \right\|_{1} + tr\left[(\Lambda^{k})^{T}\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\} \]

\vspace{1cm}

\begin{align*}
  \partial&\left\{ \lambda\left\| Z \right\|_{1} + tr\left[(\Lambda^{k})^{T}\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\} \\
  &= \partial\left\{ \lambda\left\| Z \right\|_{1} \right\} + \nabla_{\Omega}\left\{ tr\left[(\Lambda^{k})^{T}\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\} \\
  &= Sign(Z)\lambda - \Lambda^{k} - \rho\left( A\Omega^{k + 1}B - Z - C \right)
\end{align*}

where $Sign(Z)$ is the elementwise Sign operator. By setting the gradient/sub-differential equal to zero, we arrive at the following equivalence:

\[ Z_{ij} = \frac{1}{\rho}\left( \rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k} - Sign(Z_{ij})\lambda \right) \]

for all $i = 1,..., p$ and $j = 1,..., p$. We observe two scenarios:

- If $Z_{ij} > 0$ then

\[ \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} > \lambda\alpha \]

- If $Z_{ij} < 0$ then

\[ \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} < -\lambda\alpha \]

This implies that $Sign(Z_{ij}) = Sign\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}\right)$. Putting all the pieces together, we arrive at

\begin{align*}
Z_{ij}^{k + 1} &= \frac{1}{\rho}Sign\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}\right)\left( \left| \rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k} \right| - \lambda \right)_{+} \\
&= \frac{1}{\rho}Soft\left(\rho(A\Omega_{ij}^{k + 1}B - C) + \Lambda_{ij}^{k}, \lambda\right)
\end{align*}

where $Soft$ is the soft-thresholding function.

<br>\vspace{1cm}

## Stopping Criterion

In discussing the optimality conditions and stopping criterion, we will follow the steps outlined in @boyd2011distributed and cater them to the SCPME method.

Below we have three optimality conditions:

1. Primal:

\[ A\Omega^{k + 1}B - Z^{k + 1} - C = 0 \]

2. Dual:

\begin{align*}
  0 &\in \partial f\left(\Omega^{k + 1}\right) + \frac{1}{2}\left(B(\Lambda^{k + 1})^{T}A + A^{T}\Lambda^{k + 1}B^{T} \right) \\
  0 &\in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1}
\end{align*}

The first dual optimality condition is a result of taking the sub-differential of $L_{\rho}\left( \Omega^{k + 1}, Z^{k + 1}, \Lambda^{k + 1} \right)$ with respect to $\Omega^{k + 1}$ (note that we must honor the symmetric constraint of $\Omega^{k + 1}$).

We will define the left-hand side of the primal optimality condition as the primal residual: $r^{k + 1} = A\Omega^{k + 1}B - Z^{k + 1} - C$. At convergence, the optimality conditions require that $r^{k + 1} \approx 0$. The second residual we will define is the dual residual:

\[ s^{k + 1} = \frac{\rho}{2}\left( B(Z^{k + 1} - Z^{k})^{T}A + A^{T}(Z^{k + 1} - Z^{k})B^{T} \right) \]

This residual is defined from the following:

Because $\Omega^{k + 1}$ is the argument that minimizes $L_{p}\left( \Omega, Z^{k}, \Lambda^{k} \right)$,

\begin{align*}
  0 &\in \partial \left\{ f\left(\Omega^{k + 1}\right) + tr\left[ \Lambda^{k}\left( A\Omega^{k + 1}B - Z^{k} - C \right) \right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z^{k} - C \right\|_{F}^{2} \right\} \\
  &= \partial f\left(\Omega^{k + 1} \right) + \frac{1}{2}\left(B(\Lambda^{k})^{T}A + A^{T}\Lambda^{k}B^{T} \right) + \frac{\rho}{2}\left( BB^{T}\Omega^{k + 1}A^{T}A + A^{T}A\Omega^{k + 1}BB^{T} \right) \\
  &- \frac{\rho}{2}\left( A^{T}(Z^{k} + C)B^{T} + B(Z^{k} + C)^{T}A \right) \\
  &= \partial f\left(\Omega^{k + 1} \right) + \frac{1}{2}\left(B(\Lambda^{k})^{T}A + A^{T}\Lambda^{k}B^{T} \right) \\
  &+ \frac{\rho}{2}\left( B(B^{T}\Omega^{k + 1}A^{T} - (Z^{k})^{T} - C^{T})A + A^{T}(A\Omega^{k + 1}B - Z^{k} - C)B^{T} \right) \\
  &= \partial f\left(\Omega^{k + 1} \right) + \frac{1}{2}\left( B(\Lambda^{k})^{T}A + A^{T}\Lambda^{k}B^{T} \right) + \frac{\rho}{2}\left(A^{T}(A\Omega^{k + 1}B - Z^{k + 1} + Z^{k + 1} - Z^{k} - C)B^{T} \right) \\
  &+ \frac{\rho}{2}\left(B(B^{T}\Omega^{k + 1}A^{T} - (Z^{k + 1})^{T} + (Z^{k + 1})^{T} - (Z^{k})^{T} - C^{T})A \right) \\
  &= \partial f\left(\Omega^{k + 1} \right) + \frac{1}{2}\left[ B\left((\Lambda^{k})^{T} + \rho(B^{T}\Omega^{k + 1}A^{T} - (Z^{k + 1})^{T} - C^{T}) \right)A \right] \\
  &+ \frac{1}{2}\left[ A^{T}\left(\Lambda^{k} + \rho(A\Omega^{k + 1}B - Z^{k + 1} - c)B \right)B^{T} \right] + \frac{\rho}{2}\left(B(Z^{k + 1} - Z^{k})^{T}A + A^{T}(Z^{k + 1} - Z^{k})B^{T} \right) \\
  &= \partial f\left(\Omega^{k + 1} \right) + \frac{1}{2}\left(B(\Lambda^{k + 1})^{T}A + A^{T}\Lambda^{k + 1}B^{T} \right) + \frac{\rho}{2}\left(B(Z^{k + 1} - Z^{k})^{T}A + A^{T}(Z^{k + 1} - Z^{k})B^{T} \right) \\
  \Rightarrow 0 &\in \frac{\rho}{2}\left( B(Z^{k + 1} - Z^{k})^{T}A + A^{T}(Z^{k + 1} - Z^{k})B^{T} \right)
\end{align*}

Like the primal residual, at convergence the optimality conditions require that $s^{k + 1} \approx 0$. Note that the second dual optimality condition is always satisfied:

\begin{align*}
  0 &\in \partial \left\{ g\left(Z^{k + 1}\right) + tr\left[ \Lambda^{k}\left( A\Omega^{k + 1}B - Z^{k + 1} - C \right) \right] + \rho\left\| A\Omega^{k + 1}B - Z^{k + 1} - C \right\|_{F}^{2} \right\} \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k} - \rho\left(A\Omega^{k + 1}B - Z^{k + 1} - C \right) \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
\end{align*}

One possible stopping criterion is to set $\epsilon^{rel} = \epsilon^{abs} = 10^{-3}$ and stop the algorithm when $\epsilon^{pri} \leq \left\| r^{k + 1} \right\|_{F}$ and $\epsilon^{dual} \leq \left\| s^{k + 1} \right\|_{F}$ where

\begin{align*}
  \epsilon^{pri} &= \sqrt{nr}\epsilon^{abs} + \epsilon^{rel}\max\left\{ \left\| A\Omega^{k + 1}B \right\|_{F}, \left\| Z^{k + 1} \right\|_{F}, \left\| C \right\|_{F} \right\} \\
  \epsilon^{dual} &= p\epsilon^{abs} + \epsilon^{rel}\left\| \left( B(\Lambda^{k + 1})^{T}A + A^{T}\Lambda^{k + 1}B^{T} \right)/2 \right\|_{F}
\end{align*}