### 2018-07-17 07-SCPME-Connection.Rmd

# `SCPME` Connection to Joint Log-likelihood

The general optimization problem outlined in @molstad2017shrinking is to estimate $\hat{\Omega}$ such that

\begin{align}
  \hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr(S\Omega) - \log\left| \Omega \right| + \lambda\left\| A\Omega B - C \right\|_{1} \right\}
\end{align}

where $S$ is the sample covariance of $X \in \mathbb{R}^{p}$ (denominator $n$) and $\left\| \cdot \right\|_{1}$ denotes the $L_{1}$-norm. In addition, $A \in \mathbb{R}^{n \times p}, B \in \mathbb{R}^{p \times r}, \mbox{ and } C \in \mathbb{R}^{n \times r}$ are matrices assumed to be known and specified by the user. This objective function is the log-likelihood for the marginal precision matrix of $X$, which we denote here as $\Omega \equiv \Sigma_{xx}^{-1}$, plus a penalty term.

\[ \hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ l\left( \Omega \right) + P_{\lambda}\left( \Omega \right) \right\} \]

where $P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1}$. Under this particular likelihood, it is assumed that each observation of $X$ follows a $p$-variate normal distribution (ie: $X_{i} \sim N_{p}\left( \mu_{x}, \Omega^{-1} \right)$) -- though Dennis Cook and others have shown this objective function to work well as a generic loss even when the assumptions are not met.

This type of formulation leads to many new, interesting, and novel estimators for $\Omega$. For instance, we might set $A = I_{p}, B = \Sigma_{xy}, \mbox{ and } C = 0,$ in which case

\[P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\Sigma_{xy} \right\|_{1} = \lambda\left\| \beta \right\|_{1} \]

That is, this objective function estimates an $\Omega$ with respect to the marginal likelihood of $X$ under the assumption that $\beta$ is sparse (recall that $\beta \equiv \Omega\Sigma_{xy}$). We could also take $B = \left[ \Sigma_{xy}, I_{p} \right]$ so that the identity matrix (dimension $p$) is appended to the covariance matrix of $X$ and $Y$.

\[ P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\left[\Sigma_{xy}, I_{p}\right] \right\|_{1} = \lambda\left\| \beta \right\|_{1} + \lambda\left\| \Omega \right\|_{1} \]

In this case, not only are we assuming that $\beta$ is sparse, but we are also assuming sparsity in $\Omega$. Further, if we let $\mathbb{X} \in \mathbb{R}^{n \times p}$ be a matrix with rows $X_{i} - \bar{X}$ and $\mathbb{Y} \in \mathbb{R}^{n \times r}$ be a matrix with rows $Y_{i} - \bar{Y}$, we might also take $A = \mathbb{X}, B = \Sigma_{xy}, \mbox{ and } C = \mathbb{Y}$ so that

\begin{align}
  \hat{\Omega} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr(S\Omega) - \log\left| \Omega \right| + \lambda\left\| \mathbb{X}\Omega \Sigma_{xy} - \mathbb{Y} \right\|_{1} \right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr(S\Omega) - \log\left| \Omega \right| + \lambda\left\| \mathbb{Y} - \mathbb{X}\beta \right\|_{1} \right\} \nonumber
\end{align}

This estimator encourages the predicted values $\hat{\mathbb{Y}} = \mathbb{X}\hat{\beta} = \mathbb{X}\hat{\Omega}\Sigma_{xy}$ to be exactly equal to the observed responses $\mathbb{Y}$. We will name this estimator the *lasso SCPME estimator*. Note that in practice - for any of these estimators - we do not know the true covariance $\Sigma_{xy}$ but we could provide the sample estimate $\hat{\Sigma}_{xy} = \mathbb{X}^{T}\mathbb{Y}/n$.

All of these estimators fall into the SCPME (shrinking characteristics of precision matrix estimators) framework outlined in the paper. The augmented ADMM algorithm is the following:

Set $k = 0$ and repeat steps 1-6 until convergence.

1. Compute $G_{k} = \rho A^{T}\left( A\Omega_{k} B - Z_{k} - C + \rho^{-1}Y_{k} \right)$

2. Decompose $S + \left( G_{k} + G_{k}^{T} \right)/2 - \rho\tau\Omega_{k} = VQV^{T}$ (via the spectral decomposition).

3. Set $\Omega_{k + 1} = V\left( -Q + (Q^{2} + 4\rho\tau I_{p})^{1/2} \right)V^{T}/(2\rho\tau)$

4. Set $Z_{k + 1} = \mbox{soft}\left( A\Omega_{k + 1}B - C + \rho^{-1}Y_{k}, \rho^{-1}\lambda \right)$

5. Set $Y_{k + 1} = \rho\left( A\Omega_{k + 1} B - Z_{k + 1} - C \right)$

6. Replace $k$ with $k + 1$.

<br>
\vspace{1cm}

## Connection to joint log-likelihood

The last estimator in the previous section reveals an interesting connection to the joint log-likelihood of $X$ and $Y$ when we consider its analogous *ridge* version - the *ridge SCPME estimator*:

\begin{align}
  \hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr(S\Omega) - \log\left| \Omega \right| + \frac{\lambda}{2}\left\| \mathbb{X}\Omega \Sigma_{xy} - \mathbb{Y} \right\|_{F}^{2} \right\}
\end{align}

To see this, let us suppose that we have $n$ independent copies of $(Y_{i}, X_{i})$ where $Y_{i} | X_{i} \sim N_{r}\left( \mu_{y} + \beta^{T}X_{i}, \Omega_{y | x}^{-1} \right)$ and $X_{i} \sim N_{p}\left( \mu_{x}, \Omega^{-1} \right)$. That is,

\[ Y_{i} = \mu_{y} + \beta^{T}X_{i} + E_{i} \]

where $E_{i} \sim N_{r}\left( 0, \Omega_{y | x}^{-1} \right)$. Similar to before, take $\Omega_{y | x} \equiv \Sigma_{y | x}^{-1}, \Omega \equiv \Sigma_{xx}^{-1}, \mbox{ and } \beta \equiv \Omega\Sigma_{xy}$.

Below we derive the joint likelihood:

\begin{align*}
  L\left( \Omega_{y | x}, \Omega, \mu_{x}, \mu_{y}, \beta | Y, X \right) &\propto f\left(Y | X, \Omega_{y | x}, \Omega, \mu_{x}, \mu_{y}, \beta \right) \times f\left(X | \Omega, \mu_{x} \right) \\
  &= \prod_{i = 1}^{n} (2\pi)^{-r/2}\left| \Omega_{y | x} \right|^{1/2} \exp\left[ -\frac{1}{2}\left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right)^{T}\Omega_{y | x}\left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right) \right] \\
  &\times (2\pi)^{-p/2}\left| \Omega \right|^{1/2} \exp\left[ -\frac{1}{2}\left( X_{i} - \mu_{x} \right)^{T}\Omega\left( X_{i} - \mu_{x} \right) \right] \\
  &\propto \left| \Omega_{y | x} \right|^{n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right)\left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right)^{T}\Omega_{y | x} \right] \\
  &\times \left| \Omega \right|^{n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu_{x} \right)\left( X_{i} - \mu_{x} \right)^{T}\Omega \right]
\end{align*}

where $\mbox{etr}\left( \cdot \right)$ denotes the exponential trace operator. This implies that the joint log-likelihood is of the following form:

\begin{align*}
  l\left( \Omega_{y | x}, \Omega, \mu_{x}, \mu_{y}, \beta | Y, X \right) &= const. + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}\sum_{i = 1}^{n} tr\left[ \left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right)\left( Y_{i} - \mu_{y} - \beta^{T}X_{i} \right)\Omega_{y | x} \right] \\
  &+\frac{n}{2}\log\left| \Omega \right| - \frac{1}{2}\sum_{i = 1}^{n}tr\left[ \left( X_{i} - \mu_{x} \right)\left( X_{i} - \mu_{x} \right)^{T}\Omega \right]
\end{align*}

**Note:** this form simplifies significantly if we solve for $\mu_{x}$ and $\mu_{y}$ and make a few additional assumptions about our data.

\begin{align*}
 \nabla_{\mu_{x}}l\left\{\cdot\right\} &= \nabla_{\mu_{x}}\left\{ -\frac{1}{2}tr\left( -n\mu_{x}\bar{X}^{T}\Omega - n\bar{X}\mu_{x}^{T}\Omega + n\mu_{x}\mu_{x}^{T}\Omega \right) \right\} \\
 &= -\frac{1}{2}\left( -n\Omega\bar{X} - n\Omega\bar{X} + 2n\Omega\mu_{x} \right) \\
 \Rightarrow \hat{\mu_{x}} &= \bar{X} \\ \\
 \nabla_{\mu_{y}}l\left\{\cdot\right\} &= \nabla_{\mu_{y}}\left\{ -\frac{1}{2}tr\left( -n\mu_{y}(\bar{Y} - \beta^{T}\bar{X})\Omega_{y | x} - n(\bar{Y} - \beta^{T}\bar{X})\mu_{y}^{T}\Omega_{y | x} + n\mu_{y}\mu_{y}^{T}\Omega_{y | x} \right) \right\} \\
 &= -\frac{1}{2}\left( -n\Omega_{y | x}(\bar{Y} - \beta^{T}\bar{X}) - n\Omega_{y | x}(\bar{Y} - \beta^{T}\bar{X}) + 2n\Omega_{y | x}\mu_{y} \right) \\
 \Rightarrow \hat{\mu_{y}} &= \bar{Y} - \beta^{T}\bar{X}
\end{align*}

Without loss of generality, let us assume that $\bar{X} = \bar{Y} = 0$. This can be achieved by simply centering the data by the sample averages.

Putting all of this together and recalling that $\beta \equiv \Omega\Sigma_{xy}$, we have that the partially maximized joint log-likelihood is the following:

\begin{align*}
  l\left( \Omega_{y | x}, \Omega, \Sigma_{xy} | Y, X \right) &= const. + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}\sum_{i = 1}^{n} tr\left[ \left( Y_{i} - \Sigma_{xy}^{T}\Omega X_{i} \right)\left( Y_{i} - \Sigma_{xy}^{T}\Omega X_{i} \right)\Omega_{y | x} \right] \\
  &+\frac{n}{2}\log\left| \Omega \right| - \frac{1}{2}\sum_{i = 1}^{n}tr\left( X_{i}X_{i}^{T}\Omega \right) \\
  &= const. + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}tr\left[ \left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)^{T}\left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)\Omega_{y | x} \right] \\
  &+ \frac{n}{2}\log\left| \Omega \right| - \frac{1}{2}tr\left( \mathbb{X}^{T}\mathbb{X}\Omega \right) \\
  &= const. + \frac{n}{2}\log\left| \Omega_{y | x} \right| - \frac{1}{2}tr\left[ \left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)^{T}\left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{y | x} \right)\Omega_{y | x} \right] \\
  &+ \frac{n}{2}\log\left| \Omega \right| - \frac{n}{2}tr\left( S\Omega \right)
\end{align*}

where $S = \mathbb{X}^{T}\mathbb{X}/n$. Now suppose we are interested in solving for $\Omega$:

\begin{align*}
  \hat{\Omega} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}-l\left( \Omega_{y | x}, \Omega, \Sigma_{xy} | Y, X \right) \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ \frac{1}{2}tr\left[ \left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)^{T}\left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)\Omega_{y | x} \right] + \frac{n}{2}tr\left( S\Omega \right) - \frac{n}{2}\log\left| \Omega \right| \right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left( S\Omega \right) - \log\left| \Omega \right| + \frac{1}{n}\left\| \left( \mathbb{Y} - \mathbb{X}\Omega\Sigma_{xy} \right)\Omega_{y | x}^{1/2} \right\|_{F}^{2} \right\}
\end{align*}

where $\left\|\cdot\right\|_{F}$ denotes the Frobenius norm. Let us assume that $\Omega_{y | x}^{1/2} = I_{r}$ and let $\lambda = 2/n$. Then

\begin{align}
  \hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left( S\Omega \right) - \log\left| \Omega \right| + \frac{\lambda}{2}\left\| \mathbb{X}\Omega\Sigma_{xy} - \mathbb{Y} \right\|_{F}^{2} \right\}
\end{align}

Notice that this is exactly the ridge SCPME estimator. This tells us that with just a few assumptions, the ridge SCPME estimator with $A = \mathbb{X}, B = \Sigma_{xy}, \mbox{ and } C = \mathbb{Y}$ is equivalent to maximizing the joint log-likelihood of $X$ and $Y$ with respect to $\Omega$. Specifically, if we assume that, given $X$, each response is independent of one another and has equal variance (denoted as $\sigma_{y | x}^{2}$) then letting $\lambda = 2/(n\sigma_{y | x}^{2})$ and solving for the ridge SCPME estimator is equivalent to solving for the optimal $\hat{\Omega}$ under the joint log-likelihood of $X$ and $Y$.

In addition, only a very slight modification of the augmented ADMM algorithm in @molstad2017shrinking is required to compute this estimate:

Set $k = 0$ and repeat steps 1-6 until convergence.

1. Compute $G_{k} = \rho A^{T}\left( A\Omega_{k} B - Z_{k} - C + \rho^{-1}Y_{k} \right)$

2. Decompose $S + \left( G_{k} + G_{k}^{T} \right)/2 - \rho\tau\Omega_{k} = VQV^{T}$ (via the spectral decomposition).

3. Set $\Omega_{k + 1} = V\left( -Q + (Q^{2} + 4\rho\tau I_{p})^{1/2} \right)V^{T}/(2\rho\tau)$

4. Set $Z_{k + 1} = \left[ \rho\left( A\Omega_{k + 1} B - C \right) + Y_{k} \right]/(\lambda + \rho)$

5. Set $Y_{k + 1} = \rho\left( A\Omega_{k + 1} B - Z_{k + 1} - C \right)$

6. Replace $k$ with $k + 1$.

The only difference is in step 4.

There are a few takeaways we can make after revealing this connection. The first is that the estimator constructed in this section is not complete nonsense. We are, in fact, solving a very similar problem to that of optimizing the joint likelihood. The second takeaway is that the SCPME estimator offers a very efficient method of computing this joint likelihood estimator. To the best of my knowledge, no closed-form solution to the ridge SCPME estimator exists and so we must resort to some iterative algorithm like this one in order to solve it.

<br>
\vspace{1cm}

## Exploratory simulations

Let $\beta = \mathbb{B} \circ \mathbb{V}$ where $vec\left( \mathbb{B} \right) \sim N_{pr}\left( 0, I_{p} \otimes I_{r}/\sqrt{p} \right)$ and $\mathbb{V} \in \mathbb{R}^{p \times r}$ is a matrix containing $pr$ random bernoulli draws with probability 0.5 being equal to 1. The $\Sigma_{xx}$ and $\Sigma_{y | x}$ is constructed so that $\left( \Sigma_{xx} \right)_{ij} = 0.7^{\left| i - j \right|}$ and $\left( \Sigma_{y | x} \right)_{ij} = 0.7^{\left| i - j \right|}$, respectively. This ensures that the inverses will be tridiagonal (sparse). Then for $n$ independent, identically distributed samples:

\[ X_{i} \sim N_{p}\left( 0, \Sigma_{xx} \right) \]

\[ Y_{i}| X_{i} \sim N_{r}\left( \beta^{T}X_{i}, \Sigma_{y | x} \right) \]

We will take $(n = 100, p = 10, r = 5)$ for this short study and use the `data_gen` function I already have built in R.

<br>\vspace{1cm}
```{r, tidy = TRUE}

# load libraries
library(statr)
library(SCPME)
set.seed(1234)

# randomly generate the data
data = data_gen(n = 100, p = 10, r = 5)

# print Omega
Omega = qr.solve(data$SigmaX)
round(Omega, 5)

# print sample precision matrix (perhaps bad estimate)
sample = qr.solve((nrow(data$X) - 1)/nrow(data$X)*cov(data$X))
round(sample, 5)

```