### 2018-07-17 08-SCPME-Simulations.Rmd

# `SCPME` Simulations

We compare the performance of various shrinkage estimators under different data realizations. Each data scenario was replicated a total of 20 times.

<br>\vspace{1cm}

### Models

In the following estimators, let the superscript * denote the oracle estimator:

 - `OLS` = ordinary least squares estimator in *low* dimensional setting and Moore-Penrose estimator in a *high* ($p >> n$) dimensional setting

 - `ridge` = ridge regression estimator where the optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV

 - `lasso` = lasso regression estimator where the optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV

 - `oracleB` = oracle estimator with the true $\beta^{*}$ known ($\left\| Y - X\beta^{*} \right\|_{F}^{2}$)

 - `oracleO` = oracle estimator with the true $\Omega^{*}$ known ($\left\| Y - X\Omega^{*}\hat{\Sigma}_{xy} \right\|_{F}^{2}$)

 - `shrinkB` = shrinking characteristics regression estimator where optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$)

 - `shrinkBS` = shrinking characteristics regression estimator with oracle cross-covariance where optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy}^{*} \right\|_{1}$)

 - `shrinkBO` = shrinking characteristics estimator that penalizes the sum of the absolute values of $\beta = \Omega\Sigma_{xy}$ *and* $\Omega$ (penalty: $\lambda\left\| \Omega[\Sigma_{xy}, I_{p}] \right\|_{1}$)

 - `shrinkXY` = shrinking characteristics regression estimator with A = X, B = cov(X, Y), and C = Y (penalty: $\lambda\left\| X\Omega\hat{\Sigma}_{xy} - Y \right\|_{1}$)

 - `shrinkXYO` = shrinking characteristics regression estimator meant to emulate the "likelihood" for $\beta$ (using oracle $\Omega_{y | x}^{*}$) (penalty: $\lambda\left\| X\Omega\hat{\Sigma}_{xy}\Omega_{y | x}^{1/2} - Y\Omega_{y | x}^{1/2} \right\|_{1}$)

 - `glasso` = lasso penalized precision matrix where optimal tuning parameter is chosen to maximize log-likelihood via 3-fold CV.

<br>\vspace{1cm}

### Data Generation

Let $\mathbb{X} \in \mathbb{R}^{n \times p}$, $\mathbb{Y} \in \mathbb{R}^{n \times r}$, and $\beta \in \mathbb{R}^{p \times r}$ so that

\[ \mathbb{Y} = \mathbb{X}\beta + \mathbb{E} \]

These values are generated in the following way:

Let $\beta = \mathbb{B} \circ \mathbb{V}$ where $vec\left( \mathbb{B} \right) \sim N_{pr}\left( 0, I_{p} \otimes I_{r}/\sqrt{p} \right)$ and $\mathbb{V} \in \mathbb{R}^{p \times r}$ is a matrix containing $pr$ random bernoulli draws with `sparse` probability being equal to 1. In effect, each $\beta$ entry is turned off with equal probability.

Now, if `sigma = "tridiag"` then $\Sigma_{xx}$ and $\Sigma_{y | x}$ is constructed so that $\left( \Sigma_{xx} \right)_{ij} = 0.7^{\left| i - j \right|}$ and $\left( \Sigma_{y | x} \right)_{ij} = 0.7^{\left| i - j \right|}$, respectively. This ensures that the inverses will be tridiagonal (sparse).

Then for $n$ independent, identically distributed samples:

\[ X_{i} \sim N_{p}\left( 0, \Sigma_{xx} \right) \]

\[ E_{i} \sim N_{r}\left( 0, \Sigma_{y | x} \right) \]

If `sigma = "compound"` then the data is generated similarly but $\Omega_{xx} \equiv \Sigma_{xx}^{-1}$ and $\Omega_{y | x} \equiv \Sigma_{y | x}^{-1}$ both have entries equal to 1 on the diagonal and 0.9 on the off-diagonal.

An additional 1000 samples will be generated for the testing set -- regardless of the training set size ($n$). For instance, in the 3-fold cross validation procedure where `n = 100` and `p = 100`, 1000 samples would be set aside for later testing and $100*(2/3)$ samples would be used for training in each fold (ie: high dimensional setting).

Possible parameters in simulation:

 - `reps` = 20
 - `n` = (100, 1000)
 - `p` = (10, 50, 100, 150)
 - `r` = c(1, 5)
 - `sparse` = 0.5
 - `sigma` = "tridiag"

Note that prior to estimation, the intercept is removed by centering the data.

<br>\vspace{1cm}

## Plots

```{r, eval = T, echo = F}

# load data
load("images/sim5.Rdata")

# set P to numeric 
data$P %<>% as.character %>% as.numeric

# MSPE by P and R (N = 100, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + facet_wrap(~ R, scales = "free") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P and R (N = 500, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# MSPE by P (N = 100, R = 1, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 100, R == 1, sparse == 0.5, sigma == "tridiag", model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P (N = 500, R = 1, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 100, P = 150, R = 1, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# MSPE table (N = 100, sparse = 0.5, sigma = "tridiag", P = 100, R = 5)
data %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", P == 150, R == 1) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean)

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# MSPE by P (N = 100, R = 5, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 100, R == 5, sparse == 0.5, sigma == "tridiag", model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P (N = 500, R = 5, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 100, P = 150, R = 5, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# MSPE table (N = 500, sparse = 0.5, sigma = "tridiag", P = 100, R = 5)
data %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", P == 150, R == 5) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean)

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# MSPE by P and R (N = 1000, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 1000, sparse == 0.5, model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + facet_wrap(~ R, scales = "free") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P and R (N = 5000, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# MSPE by P (N = 1000, R = 1, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 1000, R == 1, sparse == 0.5, model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P (N = 5000, R = 1, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 1000, P = 150, R = 1, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# MSPE table (N = 500, sparse = 0.5, sigma = "tridiag", P = 100, R = 1)
data %>% filter(N == 1000, sparse == 0.5, sigma == "tridiag", P == 150, R == 1) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean)

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# MSPE by P (N = 1000, R = 5, sparse = 0.5, sigma = "tridiag")
data %>% filter(N == 1000, R == 5, sparse == 0.5, model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("MSPE") + ggtitle("MSPE by P (N = 5000, R = 5, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 1000, P = 150, R = 5, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# MSPE table (N = 1000, sparse = 0.5, sigma = "tridiag", P = 100, R = 5)
data %>% filter(N == 1000, sparse == 0.5, sigma == "tridiag", P == 150, R == 5) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean)

```

<br>\vspace{1cm}

## Takeaways

1. `shrinkBO` (penalty: $\lambda\left\| \Omega[\Sigma_{xy}, I_{p}] \right\|_{1}$) and `shrinkB` (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$) perform increasingly well relative (and superior) to the others when $p >> n$. I don't think it's surprising that `shrinkBO` performs the best because the embedded assumptions most closely match the data generating model.

2. It's interesting to see that in all cases `shrinkBS` (penalty: $\lambda\left\| \Omega\Sigma_{xy}^{*} \right\|_{1}$) performed worse than `shrinkB` (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$). So the takeaway is that for these two very similar estimators, the estimator with the oracle $\Sigma_{xy}$ cross-covariance matrix performs much worse than using the sample estimate $\hat{\Sigma}_{xy} = \mathbb{X}^{T}\mathbb{Y}/n$. Is this due to an error? Could the root cause be because I centered the data?

3. `OracleO` ($\left\| Y - X\Omega^{*}\hat{\Sigma}_{xy} \right\|_{F}^{2}$) performed awful in all simulations as well... possibly link to the Abundant paper?

4. `shrinkXY` (penalty: $\lambda\left\| X\Omega\hat{\Sigma}_{xy} - Y \right\|_{1}$) and `shrinkXYO` (penalty: $\lambda\left\| X\Omega\hat{\Sigma}_{xy}\Omega_{y | x}^{1/2} - Y\Omega_{y | x}^{1/2} \right\|_{1}$) did not perform well relative to the others in the high-dimensional setting. In the low-dimensional setting, they performed better but worse than `lasso` and `ridge`. But what is an appropriate data generating model for these estimators? A model that can perfectly fit the observed values? And would MSPE be the most appropriate criteria for these estimators?

<br>\vspace{1cm}

## Questions

1. Is MSPE the best criteria? Or would model error be a better metric?