### 2018-08-06 08-SCPME-Simulations.Rmd
updated format of document

# `SCPME` Simulations

## Estimator Comparison

We compare the performance of various shrinkage estimators under different data realizations. Each data scenario was replicated a total of 20 times.

<br>\vspace{1cm}

### Models

In the following estimators, let the superscript * denote the oracle estimator:

 - `OLS` = ordinary least squares estimator in *low* dimensional setting and Moore-Penrose estimator in a *high* ($p >> n$) dimensional setting

 - `ridge` = ridge regression estimator where the optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV

 - `lasso` = lasso regression estimator where the optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV

 - `oracleB` = oracle estimator with the true $\beta^{*}$ known ($\left\| Y - X\beta^{*} \right\|_{F}^{2}$)

 - `oracleO` = oracle estimator with the true $\Omega^{*}$ known ($\left\| Y - X\Omega^{*}\hat{\Sigma}_{xy} \right\|_{F}^{2}$)

 - `oracleS` = oracle estimator with the true $\Sigma_{xy}^{*}$ known and sample estimator for $\Omega$ ($\left\| Y - X\hat{\Omega}\Sigma_{xy}^{*} \right\|_{F}^{2}$)

 - `shrinkB` = shrinking characteristics regression estimator where optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$)

 - `shrinkBS` = shrinking characteristics regression estimator with oracle cross-covariance where optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy}^{*} \right\|_{1}$)

 - `shrinkBO` = shrinking characteristics estimator that penalizes the sum of the absolute values of $\beta = \Omega\Sigma_{xy}$ *and* $\Omega$ (penalty: $\lambda\left\| \Omega[\Sigma_{xy}, I_{p}] \right\|_{1}$)

 - `glasso` = lasso penalized precision matrix where optimal tuning parameter is chosen to maximize log-likelihood via 3-fold CV.

<br>\vspace{1cm}

### Data Generation

Let $\mathbb{X} \in \mathbb{R}^{n \times p}$, $\mathbb{Y} \in \mathbb{R}^{n \times r}$, and $\beta \in \mathbb{R}^{p \times r}$ so that

\[ \mathbb{Y} = \mathbb{X}\beta + \mathbb{E} \]

These values are generated in the following way:

Let $\beta = \mathbb{B} \circ \mathbb{V}$ where $vec\left( \mathbb{B} \right) \sim N_{pr}\left( 0, I_{p} \otimes I_{r}/\sqrt{p} \right)$ and $\mathbb{V} \in \mathbb{R}^{p \times r}$ is a matrix containing $pr$ random bernoulli draws with `sparse` probability being equal to 1. In effect, each $\beta$ entry is turned off with equal probability.

Now, if `sigma = "tridiag"` then $\Sigma_{xx}$ and $\Sigma_{y | x}$ is constructed so that $\left( \Sigma_{xx} \right)_{ij} = 0.7^{\left| i - j \right|}$ and $\left( \Sigma_{y | x} \right)_{ij} = 0.7^{\left| i - j \right|}$, respectively. This ensures that the inverses will be tridiagonal (sparse).

Then for $n$ independent, identically distributed samples:

\[ X_{i} \sim N_{p}\left( 0, \Sigma_{xx} \right) \]

\[ E_{i} \sim N_{r}\left( 0, \Sigma_{y | x} \right) \]

If `sigma = "compound"` then the data is generated similarly but $\Omega_{xx} \equiv \Sigma_{xx}^{-1}$ and $\Omega_{y | x} \equiv \Sigma_{y | x}^{-1}$ both have entries equal to 1 on the diagonal and 0.9 on the off-diagonal.

An additional 1000 samples will be generated for the testing set -- regardless of the training set size ($n$). For instance, in the 3-fold cross validation procedure where `n = 100` and `p = 100`, 1000 samples would be set aside for later testing and $100*(2/3)$ samples would be used for training in each fold (ie: high dimensional setting).

Possible parameters in simulation:

 - `reps` = 20
 - `n` = (100, 1000)
 - `p` = (10, 50, 100, 150)
 - `r` = c(1, 5, 10)
 - `sparse` = c(0.5)
 - `sigma` = c("tridiag")

Note that prior to estimation, the intercept is removed by centering the data.

<br>\vspace{1cm}

### Plots

#### High Dimension

Note that we are now displaying model error (ME) as opposed to mean squared prediction error (MSPE). Also, `OLS` and `oracleS` will not be show in the high dimensional setting plots because of poor performance.

```{r, eval = T, echo = F}

# load data
load("images/sim6.Rdata")

# set P to numeric 
sim6$P %<>% as.character %>% as.numeric

# ME by P and R (N = 100, sparse = 0.5, sigma = "tridiag")
sim6 %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", !model %in% c("OLS", "oracleS"), metric == "ME") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + facet_wrap(~ R, scales = "free") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("ME") + ggtitle("ME by P and R (N = 100, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

```{r, eval = T, echo = F, fig.cap = "Figure caption here."}

# ME by P (N = 100, R = 10, sparse = 0.5, sigma = "tridiag")
sim6 %>% filter(N == 100, R == 10, sparse == 0.5, sigma == "tridiag", !model %in% c("OLS", "oracleS")) %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("ME") + ggtitle("ME by P (N = 100, R = 10, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 100, P = 150, R = 10, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# ME table (N = 100, sparse = 0.5, sigma = "tridiag", P = 150, R = 10)
sim6 %>% filter(N == 100, sparse == 0.5, sigma == "tridiag", P == 150, R == 10) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean) %>% pander(caption = "Table caption here.")

```

<br>\vspace{1cm}

#### Low Dimension

```{r, eval = T, echo = F}

# ME by P and R (N = 1000, sparse = 0.5, sigma = "tridiag")
sim6 %>% filter(N == 1000, sparse == 0.5, model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + facet_wrap(~ R, scales = "free") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("ME") + ggtitle("ME by P and R (N = 1000, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# ME by P (N = 1000, R = 5, sparse = 0.5, sigma = "tridiag")
sim6 %>% filter(N == 1000, R == 5, sparse == 0.5, model != "OLS") %>% group_by(N, P, R, sparse, sigma, model) %>% summarise(mean = mean(Error)) %>% ggplot(aes(P, mean, color = model, linetype = model)) + geom_line() + scale_colour_discrete("") + scale_linetype_manual("", values = rep(c(5,6), 5)) + theme_minimal() + theme(legend.position = "bottom") + xlab("P") + labs(color = "Model") + scale_x_continuous(breaks = c(10, 50, 100, 150)) + ylab("ME") + ggtitle("ME by P (N = 1000, R = 5, sparse = 0.5, sigma = tridiag)")

```

<br>\vspace{1cm}

**Notable: N = 1000, P = 150, R = 5, sparse = 0.5, sigma = "tridiag"**

```{r, eval = T, echo = F}

# ME table (N = 1000, sparse = 0.5, sigma = "tridiag", P = 150, R = 5)
sim6 %>% filter(N == 1000, sparse == 0.5, sigma == "tridiag", P == 150, R == 5) %>% group_by(model) %>% summarise(mean = mean(Error), sd = sd(Error)) %>% arrange(mean) %>% pander(caption = "Table caption here.")

```

<br>\vspace{1cm}

### Takeaways

1. In high dimensional settings, `shrinkBO` (penalty: $\lambda\left\| \Omega[\hat{\Sigma}_{xy}, I_{p}] \right\|_{1}$) and `shrinkB` (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$) perform increasingly well relative (and superior) to the others when $p >> n$. It's not surprising that `shrinkBO` performs the best because the embedded assumptions most closely match the data generating model.

2. In the high dimension setting, the oracle estimators `oracleO` (penalty: $\lambda\left\| \Omega^{*}\hat{\Sigma}_{xy} \right\|_{1}$) and `oracleS` (penalty: $\lambda\left\| \hat{\Omega}\Sigma_{xy}^{*} \right\|_{1}$) performed worse or comparable to OLS. The poor performance of `oracleS` is likely due to the fact that the sample estimate of $\Omega$ is not identifiable when $p > n$.

3. When $n > p$ we see that `shrinkB` performs terribly. Interestingly, `shrinkBO` is still one of the better performing estimators -- though worse than both `lasso` and `ridge`.

4. It is interesting that `oracleO` performed worse than `oracleS` in the final table. It appears that estimating the covariance matrix $\Sigma_{xy}$ well is *more* important than estimating $\Omega$ well.

<br>\vspace{1cm}

## Shrinking Covariance

We compare the performance of the `shrinkB` and `shrinkBO` estimators when manually shrinking the sample covariance matrix of $X$ and $Y$ (denoted $\Sigma_{xy}$). Each data scenario was replicated a total of 20 times.

<br>\vspace{1cm}

### Models

In the following estimators, let the superscript * denote the oracle estimator:

 - `shrinkB` = shrinking characteristics regression estimator where optimal tuning parameter is chosen to minimize in-sample MSPE via 3-fold CV (penalty: $\lambda\left\| \Omega\hat{\Sigma}_{xy} \right\|_{1}$)

 - `shrinkBO` = shrinking characteristics estimator that penalizes the sum of the absolute values of $\beta = \Omega\Sigma_{xy}$ *and* $\Omega$ (penalty: $\lambda\left\| \Omega[\Sigma_{xy}, I_{p}] \right\|_{1}$)

For each of the estimators, $\Sigma_{xy}$ is estimated by the sample covariance matrix $\hat{\Sigma}_{xy}$ times some constant factor $k \in (0.1, 0.2, ..., 0.9, 1)$. We are shrinking $\hat{\Sigma}_{xy}$ by a constant to determine if there are any benefits to doing so.

<br>\vspace{1cm}

### Data Generation

Let $\mathbb{X} \in \mathbb{R}^{n \times p}$, $\mathbb{Y} \in \mathbb{R}^{n \times r}$, and $\beta \in \mathbb{R}^{p \times r}$ so that

\[ \mathbb{Y} = \mathbb{X}\beta + \mathbb{E} \]

These values are generated in the following way:

Let $\beta = \mathbb{B} \circ \mathbb{V}$ where $vec\left( \mathbb{B} \right) \sim N_{pr}\left( 0, I_{p} \otimes I_{r}/\sqrt{p} \right)$ and $\mathbb{V} \in \mathbb{R}^{p \times r}$ is a matrix containing $pr$ random bernoulli draws with `sparse` probability being equal to 1. In effect, each $\beta$ entry is turned off with equal probability.

Now, if `sigma = "tridiag"` then $\Sigma_{xx}$ and $\Sigma_{y | x}$ is constructed so that $\left( \Sigma_{xx} \right)_{ij} = 0.7^{\left| i - j \right|}$ and $\left( \Sigma_{y | x} \right)_{ij} = 0.7^{\left| i - j \right|}$, respectively. This ensures that the inverses will be tridiagonal (sparse).

Then for $n$ independent, identically distributed samples:

\[ X_{i} \sim N_{p}\left( 0, \Sigma_{xx} \right) \]

\[ E_{i} \sim N_{r}\left( 0, \Sigma_{y | x} \right) \]

If `sigma = "compound"` then the data is generated similarly but $\Omega_{xx} \equiv \Sigma_{xx}^{-1}$ and $\Omega_{y | x} \equiv \Sigma_{y | x}^{-1}$ both have entries equal to 1 on the diagonal and 0.9 on the off-diagonal.

An additional 1000 samples will be generated for the testing set -- regardless of the training set size ($n$). For instance, in the 3-fold cross validation procedure where `n = 100` and `p = 100`, 1000 samples would be set aside for later testing and $100*(2/3)$ samples would be used for training in each fold (ie: high dimensional setting).

Possible parameters in simulation:

 - `reps` = 20
 - `n` = (100, 1000)
 - `p` = (10, 50, 100, 150, 200)
 - `r` = c(1, 5, 10)
 - `sparse` = c(0.5)
 - `sigma` = c("tridiag", "compound")

Note that prior to estimation, the intercept is removed by centering the data.

<br>\vspace{1cm}

### Plots

```{r, eval = T, echo = F}

# load data
load("images/sigma.Rdata")

# design color palette
bluetowhite <- c("#000E29", "white")

# change to numeric
sigma$lam %<>% as.character %>% as.numeric
sigma$const %<>% as.character %>% as.numeric

```

#### High Dimension

##### `shrinkB`

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkB MSPE by lam and const (N = 100, P = 200, R = 10, sigma = "tridiag")
d = sigma %>% filter(N == 100, P == 200, R == 10, sparse == 0.5, sigma == "tridiag", model == "shrinkB") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkB MSPE (N = 100, P = 200, R = 10, sigma = tridiag)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkB MSPE by lam and const (N = 100, P = 200, R = 10, sigma = "compound")
d = sigma %>% filter(N == 100, P == 200, R == 10, sparse == 0.5, sigma == "compound", model == "shrinkB") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkB MSPE (N = 100, P = 200, R = 10, sigma = compound)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

##### `shrinkBO`

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkBO MSPE by lam and const (N = 100, P = 200, R = 10, sigma = "tridiag")
d = sigma %>% filter(N == 100, P == 200, R == 10, sparse == 0.5, sigma == "tridiag", model == "shrinkBO") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkBO MSPE (N = 100, P = 200, R = 10, sigma = tridiag)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkBO MSPE by lam and const (N = 100, P = 200, R = 10, sigma = "compound")
d = sigma %>% filter(N == 100, P == 200, R == 10, sparse == 0.5, sigma == "compound", model == "shrinkBO") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkBO MSPE (N = 100, P = 200, R = 10, sigma = compound)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

#### Low Dimension

##### `shrinkB`

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkB MSPE by lam and const (N = 1000, P = 50, R = 1, sigma = "tridiag")
d = sigma %>% filter(N == 1000, P == 50, R == 1, sparse == 0.5, sigma == "tridiag", model == "shrinkB") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkB MSPE (N = 1000, P = 50, R = 1, sigma = tridiag)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkB MSPE by lam and const (N = 1000, P = 50, R = 1, sigma = "compound")
d = sigma %>% filter(N == 1000, P == 50, R == 1, sparse == 0.5, sigma == "compound", model == "shrinkB") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkB MSPE (N = 1000, P = 50, R = 1, sigma = compound)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

#### `shrinkBO`

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkBO MSPE by lam and const (N = 1000, P = 10, R = 1, sigma = "tridiag")
d = sigma %>% filter(N == 1000, P == 50, R == 1, sparse == 0.5, sigma == "tridiag", model == "shrinkBO") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkBO MSPE (N = 1000, P = 50, R = 1, sigma = tridiag)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

```{r, eval = T, echo = F}

# shrinkBO MSPE by lam and const (N = 1000, P = 50, R = 1, sigma = "compound")
d = sigma %>% filter(N == 1000, P == 50, R == 1, sparse == 0.5, sigma == "compound", model == "shrinkBO") %>% group_by(lam, const) %>% summarise(mean = mean(Error)) %>% mutate(augmean = 1/(c(mean)))

min = d[d$mean == min(d$mean),]

d %>% ggplot(aes(const, log10(lam))) + geom_raster(aes(fill = augmean)) + scale_fill_gradientn(colours = colorRampPalette(bluetowhite)(2), guide = "none") + theme_minimal() + labs(title = "shrinkBO MSPE (N = 1000, P = 50, R = 1, sigma = compound)", caption = paste("**Optimal: log10(lam) = ", round(log10(min$lam), 3), ", const = ", round(min$const, 3), sep = ""))

```

<br>\vspace{1cm}

### Takeaways

1. In the high dimensional settings, it does appear that shrinking the sample covariance matrix $\hat{\Sigma}_{xy}$ by a constant factor helps in reducing MSPE. This is indicated by the fact that the optimal `const` hovers around 0.5 for each of the estimators `shrinkB` and `shrinkBO`.

2. In the low dimension settings, it appears that shrinking the covariance matrix is not beneficial.

3. The type of oracle precision matrix (tridiagonal or compound symmetric) does not appear to be too influential with regards to the affect of shrinking the covariance matrix.

4. The larger the dimension of $Y$, the more beneficial shrinkage will likely be (not that surprising)