### 2019-01-09 05-SCPME-Package.Rmd

# `SCPME` R Package

`SCPME` is an implementation of the proposed augmented ADMM algorithm in @molstad2017shrinking for solving the previous optimization problem. In addition, this package places a big emphasis on flexibility that allows for rapid experimentation for the end user.

We will illustrate this with a short simulation and show some of the new and interesting estimators for $\Omega$ that are a result of this penalty.

<br>\vspace{0.5cm}

## Simulation

Let us generate some data. For this example, we will assume

\[ Y_{i} = \mu_{y} + \beta^{T}\left(X_{i} - \mu_{x}\right) + E_{i} \]

where $E_{i} \sim N_{r}\left( 0, \Omega_{y | x}^{-1} \right)$, $X_{i} \sim N_{p}\left( \mu_{x}, \Omega^{-1} \right)$ and we are interested in estimating the marginal precision matrix of $X$ (denoted $\Omega$).

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}
library(SCPME)
set.seed(123)

# generate data from a sparse oracle precision matrix.
# we can use the built-in `data_gen` function

# generate 100 x 5 X data matrix and 100 x 1 Y data matrix
data = data_gen(p = 5, n = 100, r = 1)

# the default regression coefficients are sparse
data$betas

# default oracle precision matrix is also sparse
round(qr.solve(data$SigmaX), 5)

# snap shot of X data matrix
head(data$X)

# snap shot of Y data matrix
head(data$Y)

```
<br>\vspace{0.5cm}

We have generated 100 samples of the random variable $X \in \mathbb{R}^{5}$ and 100 samples of the random variable $Y \in \mathbb{R}$. It turns out that this particular oracle covariance matrix for $X$ (tapered matrix) has an inverse that is sparse (tri-diagonal). That is, the precision matrix has many zeros.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# print oracle covariance matrix
data$SigmaX

# print inverse covariance matrix (omega)
round(qr.solve(data$SigmaX), 5)

```
<br>\vspace{0.5cm}

In this particular setting, we could estimate $\Omega$ by taking the inverse of the sample covariance matrix $\hat{S} = \sum_{i = 1}^{n}(X_{i} - \bar{X})(X_{i} - \bar{X})^{T}/n$:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# print inverse of sample precision matrix (perhaps a bad estimate)
round(qr.solve(cov(data$X)*(nrow(data$X) - 1)/nrow(data$X)), 5)

```
<br>\vspace{0.5cm}

However, because $\Omega$ is sparse, this estimator will likely perform very poorly. Notice the number of zeros in our oracle precision matrix compared to the inverse of the sample covariance matrix. Instead, we will use `SCPME` to estimate $\Omega$.

By default, `SCPME` will estimate $\Omega$ using a lasso penalty ($A = I_{p}, B = I_{p}, \mbox{ and } C = 0$) and choose the optimal `lam` tuning parameter that minimizes the mean squared prediction error for the regression of the variable $Y$ on $X$.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# lasso penalty
shrink(X = data$X, Y = data$Y)

```
<br>\vspace{0.5cm}

However, we could also select the optimal tuning parameter based on other criteria, such as log-likelihood. Other options include: AIC, BIC, and penalized log-likelihood (`penloglik`).

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# lasso penalty with crit.cv = loglik
shrink(X = data$X, Y = data$Y, crit.cv = "loglik")

```
<br>\vspace{0.5cm}

`SCPME` also has the capability to provide plots for the cross validation errors. In the heatmap plot below, the more bright (white) areas of the heat map correspond to a better tuning parameter selection.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# produce CV heat map
shrink = shrink(X = data$X, nlam = 50, crit.cv = "BIC")
plot(shrink, type = "heatmap")

```
<br>\vspace{0.5cm}

Note that in the previous plot, it is not necessary to provide the $Y$ data matrix because neither the penalty nor the cross validation criteria depends on the values of $Y$.

We can also produce a line graph of the cross validation errors:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# produce line graph
plot(shrink, type = "line")

```
<br>\vspace{0.5cm}

We also have the option to print *all* of the estimated precision matrices for each tuning parameter using the `path` option. This option should be used with *extreme* care when the dimension and sample size is large -- you may run into memory issues.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# keep all estimates using path
shrink = shrink(X = data$X, crit.cv = "loglik", path = TRUE)

# print only first three objects
shrink$Path[,,1:3]

```
<br>\vspace{0.5cm}

Recall that all of the estimators so far have used a lasso penalty that penalizes the sum of the absolute value of all the entries in $\Omega$. In effect, this penalty embeds an assumption in our estimate that the true $\Omega$ is sparse.

The flexibility of the penalty described in @molstad2017shrinking allows us to make other assumptions as well. For instance, in the penalty we could set $A = I_{p}, B = \Sigma_{xy}$ where $\Sigma_{xy}$ is the covariance matrix of $X$ and $Y$, and $C = 0$. In which case

\[P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\Sigma_{xy} \right\|_{1} = \lambda\left\| \beta \right\|_{1} \]

This objective function estimates an $\Omega$ via the marginal log-likelihood of $X$ under the assumption that the forward regression coefficient $\beta$ is sparse (recall that $\beta \equiv \Omega\Sigma_{xy}$). Of course, we do not know the true covariance matrix $\Sigma_{xy}$ but we could use the sample estimate instead.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# assume sparsity in beta
lam_max = max(abs(crossprod(data$X, data$Y)))
(shrink = shrink(X = data$X, Y = data$Y, B = cov(data$X, data$Y), lam.max = lam_max, nlam = 20))

# plot CV errors
plot(shrink)

```
<br>\vspace{0.5cm}

Note that we specified the maximum `lam` value in the previous function to expand the tuning parameter grid.

Conveniently, with these settings, the augmented ADMM algorithm also solves for the estimated $\beta$ coefficient matrix simultaneously:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# print estimated beta matrix
shrink$Z

```
<br>\vspace{0.5cm}

Another possible penalty is to set $B = \left[ \Sigma_{xy}, I_{p} \right]$ so that the identity matrix (dimension $p$) is appended to the covariance matrix of $X$ and $Y$.

\[ P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\left[\Sigma_{xy}, I_{p}\right] \right\|_{1} = \lambda\left\| \beta \right\|_{1} + \lambda\left\| \Omega \right\|_{1} \]

In this case, not only are we assuming that $\beta$ is sparse, but we are also assuming sparsity in $\Omega$.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# assume sparsity in beta AND omega
(shrink = shrink(X = data$X, Y = data$Y, B = cbind(cov(data$X, data$Y), diag(ncol(data$X))), lam.max = 10, lam.min.ratio = 1e-4, nlam = 20))

# print estimated beta
shrink$Z[, 1, drop = FALSE]

# plot CV errors
plot(shrink)

```
<br>\vspace{0.5cm}

## More advanced options

A huge issue in precision matrix estimation is the computational complexity when the sample size and dimension of our data is particularly large. There are a number of built-in options in `SCPME` that can be used to improve computation speed:

 - Reduce the number of `lam` values during cross validation. The default number is 10.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# reduce number of lam to 5
shrink = shrink(X = data$X, Y = data$Y, nlam = 5)

```
<br>\vspace{0.5cm}

 - Reduce the number of `K` folds during cross validation. The default number is 5.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# reduce number of folds to 3
shrink = shrink(X = data$X, Y = data$Y, K = 3)

```
<br>\vspace{0.5cm}

 - Relax the convergence critera for the ADMM algorithm using the `tol.abs` and `tol.rel` options. The default for each is 1e-4.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# relax convergence criteria
shrink = shrink(X = data$X, Y = data$Y, tol.abs = 1e-3, tol.rel = 1e-3)

```
<br>\vspace{0.5cm}

 - Adjust the maximum number of iterations. The default is 1e4.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# adjust maximum number of iterations
shrink = shrink(X = data$X, Y = data$Y, maxit = 1e3)

```
<br>\vspace{0.5cm}

 - Adjust `adjmaxit`. This allows the user to adjust the maximum number of iterations *after* the first `lam` tuning parameter has fully converged during cross validation. This allows for *one-step estimators* and can greatly reduce the time required for the cross validation procedure while still choosing near-optimal tuning parameters.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# adjust adjmaxit
shrink = shrink(X = data$X, Y = data$Y, adjmaxit = 2)

```
<br>\vspace{0.5cm}

 - We can also opt to run our cross validation procedure in parallel. The user should check how many cores are on their system before using this option

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE, eval = FALSE}

# parallel CV
shrink = shrink(X = data$X, Y = data$Y, cores = 2)

```
<br>\vspace{0.5cm}