### 2019-01-14 01-intro.Rmd
finished modifying structure of appendix

# Introduction {#intro}

In many statistical applications, estimating the covariance matrix, $\Sigma$, for a set of random variables is a critical task. The covariance matrix is useful because it characterizes the *relationship* between variables. For instance, suppose we have three variables $X, Y, \mbox{ and } Z$ and their covariance matrix is of the form

\begin{equation}
\Sigma_{xyz} = \begin{pmatrix}
1 & 0 & 0.5 \\ 
0 & 1 & 0 \\ 
0.5 & 0 & 1
\end{pmatrix}
(\#eq:ex)\notag
\end{equation}

We can gather valuable information from this matrix. First of all, we know that each of the variables has an equal variance of 1. Second,  we know that variables $X$ and $Y$ are likely independent because the covariance between the two is equal to 0. This implies that any information in $X$ is useless in trying to gather information about $Y$. Lastly, we know that variables $X$ and $Z$ are moderately, positively correlated because their covariance is 0.5.

Unfortunately, estimating $\Sigma$ well is often computationally expensive and, in a few settings, extremely challenging. For this reason, emphasis in the literature and elsewhere has been placed on estimating the inverse of $\Sigma$ which we like to denote as $\Omega \equiv \Sigma^{-1}$.

If we have data that is normally distributed with mean $\mu$ and variance $\Sigma$ (that is, $X_{i} \sim N_{p}\left(\mu, \Sigma \right)$), the optimal estimator for $\Omega$ with respect to the log-likelihood is of the form

\begin{equation}
\hat{\Omega}_{MLE} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| \right\}
(\#eq:mleoptim)
\end{equation}

where $S$ denotes the usual sample estimator ($S = \sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)^{T})$). As in regression settings, we can construct *penalized* log-likelihood estimators by adding a penalty term, $P\left(\Omega\right)$, to the log-likelihood so that

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\}
(\#eq:penoptim)
\end{equation}

$P\left( \Omega \right)$ is often of the form $P\left(\Omega \right) = \lambda\|\Omega \|_{F}^{2}/2$ or $P\left(\Omega \right) = \|\Omega\|_{1}$ where $\lambda > 0$, $\left\|\cdot \right\|_{F}^{2}$ is the Frobenius norm and we define $\left\|A \right\|_{1} = \sum_{i, j} \left| A_{ij} \right|$. These penalties are the ridge and lasso, respectively. The penalty proposed in @molstad2017shrinking, however, is of the form $P\left(\Omega\right) = \left\|A\Omega B - C\right\|_{1}$ so that the general optimization problem is

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr(S\Omega) - \log\left| \Omega \right| + \lambda\left\| A\Omega B - C \right\|_{1} \right\}
(\#eq:molstadeq)
\end{equation}

`SCPME` is an implementation of the proposed augmented ADMM algorithm in @molstad2017shrinking for solving the previous optimization problem. In addition, this package places a big emphasis on flexibility that allows for rapid experimentation for the end user.

We will illustrate this with a short simulation and show some of the new and interesting estimators for $\Omega$ that are a result of this penalty.

- Why do we care?

- What is SCPME?

- Guide

- Model and notation