### 2019-03-10 01-intro.Rmd
added notations and restructured

# Introduction {#intro}

In many statistical applications, estimating the covariance matrix, $\Sigma$, for a set of random variables is a critical task. The covariance matrix is useful because it characterizes the *relationship* between variables. For instance, suppose we have three variables $X, Y, \mbox{ and } Z$ and their covariance matrix is of the form

\begin{equation}
\Sigma_{xyz} = \begin{pmatrix}
1 & 0 & 0.5 \\ 
0 & 1 & 0 \\ 
0.5 & 0 & 1
\end{pmatrix}
(\#eq:ex)\notag
\end{equation}

We can gather valuable information from this matrix. First of all, we know that each of the variables has an equal variance of 1. Second,  we know that variables $X$ and $Y$ are likely independent because the covariance between the two is equal to 0. This implies that any information in $X$ is useless in trying to gather information about $Y$. Lastly, we know that variables $X$ and $Z$ are moderately, positively correlated because their covariance is 0.5.

Unfortunately, estimating $\Sigma$ well is often computationally expensive and, in a few settings, extremely challenging. For this reason, emphasis in the literature and elsewhere has been placed on estimating the inverse of $\Sigma$ which we like to denote as $\Omega \equiv \Sigma^{-1}$.

### references

The reference we need here is @molstad2017shrinking!

1. From @molstad2017shrinking: "To fit many predictive models, only a characteristic of the population precision matrix needs to be estimated. For example, in binary linear discriminant analysis, the population precision matrix is needed for prediction only through the product of the precision matrix and the difference between the two conditional distribution mean vectors. Many authors have proposed methods that directly estimate this characteristic @cai2011direct; @fan2012road; @mai2012direct."

2. From @molstad2017shrinking: "@chen2016regularized proposed a method for estimating the characteristic $\Omega_{*}\mu_{*}$ directly, but like the direct linear discriminant methods, this apporach does not lead to an estimate of $\Omega_{*}$."

3. Discuss linear regression application.... where $\beta = \Omega_{*}\Sigma_{*XY}$. Like @witten2009covariance, this approach estimates the regression coefficients "using shrinkage estimators of the marginal precision matrix for the predictors".

- Guide

- Model and notation

## Outline

- Why do we care about covariance matrices? Covariance matrices are an object that summarizes the relationship between variables.

- What is a precision matrix? Precision matrix is the inverse of a covariance matrix.

- Precision matrices are powerful, essential in representing relationships in variables

- In fact, required to be estimated to fit many statistical models.

- That is, we must provide an object that accurately summarizes the relationship and interaction between variables in our model.

- Many proposed shrinkage estimators when p is particularly large. Shrinkage.

- However, to fit these models, Molstad and Rothman focused on the fact that onnly a "characteristic" of a precision matrix is required to fit these models. "We propose a framework to shrink a user-specified characteristic of a precision matrix estimator."

- For instance, LDA, linear regression, ...

- Their approach/methodology was published in BLAH. Shrinking characteristic of precision matrix estimators.

- BREAK

- We wanted to explore the regression application that was mentioned but not pursued/investigated

- This document will outline the research and work leading up to that

- The document is outlined as follows...

- We will begin chapter 2 with a brief introduction to precision matrix estimation. This will include mentions of estimation methods/approaches and we will spend a particular amount of time elaborating on one approach which is estimation via the ADMM algorithm. This dicussion will be useful when we circle back to SCPME, the regression application, and the augmented ADMM algorithm developed in BLAH.

- Lastly, the document will end with two brief tutorials for the R packages ADMMsigma and SCPME. These packages, developed by myself, will be mentioned throughout the document and were instrumental in this pursued line of research and facilitated many simulations and research direction explored throughout. Both packages have since been published on CRAN

### Notation and Definitions

For strictly positive integers $n$ and $p$, we will denote $\mathbb{R}^{n \times p}$ as the class of real matrices with dimenson $n \times p$. The class of real, symmetric matrices with dimension $p \times p$ will be denoted as $\mathbb{S}^{p}$ and $\mathbb{S}^{p}_{+}$ if we further require the object to be positive definite. The sample size in a given data set will most often correspond with $n$ and $p$ will correspond with the dimension of the prediction vector. If the dimension of the response vector exceeds one, we will denote it as $r$.

Most matrices will take the form of either $\Sigma$, which should be taken to represent the population covariance matrix, or $\Omega$, which should be taken to represent the population *precision* matrix. Note that the precision matrix is simply the inverse of the covariance matrix ($\Omega \equiv \Sigma^{-1}$). We will sometimes add a subscript star if the object is oracle - or known- a priori ($\Omega_{*}$) and the population object's estimator that minimizes/maximizes a pre-specified objective function will have a hat attached to it ($\hat{\Omega}$).

There will be significant matrix algebra notation throughout the document. The trace operator which sums the diagonal elements of a matrix will take the form $tr\left(\cdot\right)$ and the *exponential* trace operator will be denoted similarly as $etr\left(\cdot\right)$. The determinant of a matrix $\mathbf{A}$ will be denoted as $\left|\mathbf{A}\right|$ but may also take the form as simply $det\left(\mathbf{A}\right)$. The kronecker product of two matrices $\mathbf{A}$ and $\mathbf{B}$ will be denoted as $\mathbf{A} \otimes \mathbf{B}$. Lastly, the Frobenius norm which sums the square of all entries in a matrix will be denoted as $\left\|\mathbf{A}\right\|_{F}$ and we will define $\left\|\mathbf{A}\right\|_{1} := \sum_{i, j}\left|\mathbf{A}_{ij}\right|$ where the $i$-$j$th element in matrix $\mathbf{A}$ is denoted as $\left(\mathbf{A}\right)_{ij}$ or simply $\mathbf{A}_{ij}$.