### 2019-03-10 07-SCPME-Package.Rmd

# `SCPME` R Package

`SCPME` is an R package that estimates a penalized precision matrix via a modified alternating direction method of multipliers (ADMM) algorithm as described in @molstad2017shrinking. Specifically, the modified ADMM algorithm solves the following optimization problem:

\[ \hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S \Omega\right) - \log\det\left(\Omega\right) + \lambda\left\| A\Omega B - C \right\|_{1} \right\} \]

where $\lambda > 0$ is a tuning parameter, $A, B, \mbox{ and } C$ are known, user-specified matrices, and we define $\left\|A \right\|_{1} = \sum_{i, j} \left| A_{ij} \right|$.

This form of penalty leads to many new, interesting, and novel estimators for the precision matrix $\Omega$. Users can construct matrices $A, B, \mbox{ and } C$ so that emphasis is placed on the sum, absolute value of a *characteristic* of $\Omega$. We will explore a few of these estimators in the tutorial section.

A list of functions contained in the package can be found below:

* `shrink()` computes the estimated precision matrix

* `data_gen()` data generation function (for convenience)

* `plot.shrink()` produces a heat map or optional line graph for the cross validation errors

## Installation

```{r, eval = FALSE}
# The easiest way to install is from CRAN
install.packages("SCPME")

# You can also install the development version from GitHub:
# install.packages("devtools")
devtools::install_github("MGallow/SCPME")
```
<br>\vspace{0.5cm}

This package is hosted on Github at [github.com/MGallow/SCPME](https://github.com/MGallow/SCPME). The project website is located at [mattxgalloway.com/SCPME](http://mattxgalloway.com/SCPME/).

## Tutorial

The primary function in the `SCPME` package is `shrink()`. The input values $X$ is an $n \times p$ data matrix so that there are $n$ rows each representing an observation and $p$ columns each representing a unique variable and $Y$ is an $n \times r$ response matrix where $r$ is the dimension of the response vector. By default, `SCPME` will estimate $\Omega$ using a lasso penalty ($A = I_{p}, B = I_{p}, \mbox{ and } C = 0$) and choose the optimal $\lambda$ tuning parameter that minimizes the mean squared prediction error for the regression of the variable $Y$ on $X$ (here $I_{p}$ denotes a $p$-dimension identity matrix). If $Y$ is not provided, then tuning parameter selection will be based on the validation likelihood. Note that $\Omega$ (perhaps better denoted here as $\Omega_{x}$) will only have meaningful shrinkage unless the data vector $X \in \mathbb{R^{p}}$ is multi-dimensional ($p > 1$).

In this example, the data matrix is $100 \times 5$ and the response is generated according to the following model:

\[ Y_{i} = \beta^{T}X_{i} + E_{i} \]

where $E_{i} \sim N\left( 0, 1 \right)$ and $X_{i}$ is generated from a multivariate normal distribution with mean zero and tapered oracle covariance matrix $S$. A tapered covariance matrix has the property that its inverse - the precision matrix - is tri-diagonal. Estimating this oracle precision matrix well and efficiently will be our primary interest. In addition, $\beta$ is randomly generated and sparse. The data will be generated using the `data_gen()` function contained in the package.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}
library(SCPME)
set.seed(123)

# generate 100 x 5 X data matrix and 100 x 1 Y data matrix
data = data_gen(p = 5, n = 100, r = 1)

# the oracle regression coefficients are sparse
data$betas

```

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# shrink sum absolute entries in omega
shrink(X = data$X, Y = data$Y)

```
<br>\vspace{0.5cm}

Notice here that the estimated precision matrix is *not* sparse. This is due to the fact that our cross validation criteria is the mean-squared prediction error. We can estimate a new precision matrix using the validation likelihood as the cross validation criteria with the following command:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# shrink sum absolute entries in omega
shrink(X = data$X, Y = data$Y, crit.cv = "loglik")

```
<br>\vspace{0.5cm}

All of the estimators so far have used a lasso penalty that penalizes the sum of the absolute value of all the entries in $\Omega$ ($A = I_{p}, B = I_{p}, \mbox{ and } C = 0$). In effect, this penalty embeds an assumption in our estimate that the true population precision matrix, $\Omega$, is sparse. The flexibility of the penalty described in @molstad2017shrinking allows us to make other assumptions as well. For instance, in the penalty we could set $A = I_{p}, B = \Sigma_{xy}$ where $\Sigma_{xy}$ is the covariance matrix of $X$ and $Y$, and $C = 0$. In which case our penalty function

\[P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\Sigma_{xy} \right\|_{1} = \lambda\left\| \beta \right\|_{1} \]

This objective function estimates an $\Omega$ via the marginal log-likelihood of $X$ under the assumption that the forward regression coefficient $\beta$ is sparse (recall that $\beta \equiv \Omega\Sigma_{xy}$). Of course, in practice, we do not know the true covariance matrix $\Sigma_{xy}$ but we might consider using the sample estimate $\hat{\Sigma}_{xy} = \sum_{i = 1}^{n}\left(X_{i} - \bar{X}\right)\left(Y_{i} - \bar{Y}\right)^{T}/n$.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# assume sparsity in beta
lam_max = max(abs(crossprod(data$X, data$Y)))
(shrink = shrink(X = data$X, Y = data$Y, B = cov(data$X, data$Y), lam.max = lam_max, nlam = 20))

```

Note that we specified the maximum `lam` value in the previous function to expand the tuning parameter grid. With these settings, the augmented ADMM algorithm also solves for the estimated $\beta$ coefficient matrix simultaneously:

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# print estimated beta matrix
shrink$Z

```
<br>\vspace{0.5cm}

Another possible penalty is to set $B = \left[ \Sigma_{xy}, I_{p} \right]$ so that the identity matrix is appended to the covariance matrix of $X$ and $Y$. That is, the penalty $P$, is constructed as

\[ P_{\lambda}\left(\Omega \right) = \lambda\left\| A\Omega B - C \right\|_{1} = \lambda\left\| \Omega\left[\Sigma_{xy}, I_{p}\right] \right\|_{1} = \lambda\left\| \beta \right\|_{1} + \lambda\left\| \Omega \right\|_{1} \]

In this case, we are equally penalizing the sum, absolute values of entries in $\beta$ *and* $\Omega$ which embeds an assumption that $\beta$ and $\Omega$ are both sparse.

<br>\vspace{0.5cm}
```{r, message = FALSE, echo = TRUE}

# assume sparsity in beta AND omega
(shrink = shrink(X = data$X, Y = data$Y, B = cbind(cov(data$X, data$Y), diag(ncol(data$X))), lam.max = 10, lam.min.ratio = 1e-4, nlam = 20))

# print estimated beta
shrink$Z[, 1, drop = FALSE]

```
<br>\vspace{0.5cm}

`SCPME` also has the capability to provide plots (heatmaps and line graphs) for the cross validation errors. In the heatmap plot below, the more bright, white areas correspond to a better tuning parameter selection (lower cross validation error).

<br>\vspace{0.5cm}
```{r SCPMEpackage1, message = FALSE, echo = TRUE, fig.cap = "CV heatmap for SCPME tutorial", out.width = "80%"}

# produce CV heat map
plot(shrink, type = "heatmap")

```

```{r SCPMEpackage2, message = FALSE, echo = TRUE, fig.cap = "CV line graph for SCPME tutorial", out.width = "80%"}

# produce line graph
plot(shrink, type = "line")

```
<br>\vspace{0.5cm}

`SCPME` has a number of more advanced options including alternative convergence criteria and parallel CV that are explained in detail on the project website at [mattxgalloway.com/SCPME](http://mattxgalloway.com/SCPME/).