### 2019-03-10 02-PME.Rmd

# Precision Matrix Estimation

## Background

The foundation of much of the precision matrix estimation literature is the gaussian negative log-likelihood. We consider the case where we observe $n$ independent, identically distributed (iid) copies of the random variable ($X_{i}$) where $X_{i} \in \mathbb{R}^{p}$ is normally distributed with some mean, $\mu$, and some variance, $\Omega^{-1}$. That is, $X_{i}$ follows a $p$-dimensional normal distribution $X_{i} \sim N_{p}\left( \mu, \Omega^{-1} \right)$. Note that this formulation implies that $\mu \in \mathbb{R}^{p}$, $\Omega^{-1} \in \mathbb{S}_{+}^{p}$, and the probability of observing a single $X_{i}$ is

\begin{equation} 
  f\left(X_{i}; \mu, \Omega^{-1}\right) = (2\pi)^{-p/2}\left| \Omega \right|^{1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)'\Omega\left( X_{i} - \mu \right) \right] 
(\#eq:normallikone)\notag
\end{equation}

Furthermore, because we assume pairwise independence between each observation, we know that for a sample of size $n$, the probability of observing all $X_{1}, ..., X_{n}$ is equal to

\begin{equation} 
\begin{split}
  f(X_{1}, ..., X_{n}; \mu, \Omega^{-1}) &= \prod_{i = 1}^{n}(2\pi)^{-p/2}\left| \Omega \right|^{1/2}\exp\left[ -\frac{1}{2}\left( X_{i} - \mu \right)'\Omega\left( X_{i} - \mu \right) \right] \\
  &= (2\pi)^{-np/2}\left| \Omega \right|^{n/2}\mbox{etr}\left[ -\frac{1}{2}\sum_{i = 1}^{n}\left( X_{i} - \mu \right)\left( X_{i} - \mu \right)'\Omega \right]
\end{split}
(\#eq:normallik)\notag
\end{equation}

where $\mbox{etr}\left( \cdot \right)$ denotes the exponential trace operator. Therefore, the gaussian log-likelihood for $\mu$ and $\Omega$ can be written as

\begin{equation}
  l(\mu, \Omega | X) = const. + \frac{n}{2}\log\left| \Omega \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \mu \right)\left(X_{i} - \mu \right)'\Omega \right]
(\#eq:normalloglik)\notag
\end{equation}

It is relatively straight forward to show that the maximum likelihood estimator (MLE) for $\mu$ is $\hat{\mu}^{MLE} = \sum_{i = 1}^{n}X_{i}/n \equiv \bar{X}$, so that the partially maximized gaussian log-likelihood for $\Omega$ can be written as

\begin{equation}
\begin{split}
  l(\Omega | X) &= \frac{n}{2}\log\left| \Omega \right| - tr\left[ \frac{1}{2}\sum_{i = 1}^{n}\left(X_{i} - \bar{X} \right)\left(X_{i} - \bar{X} \right)'\Omega \right] \\
  &= \frac{n}{2}\log\left| \Omega \right| - \frac{n}{2}tr\left( S\Omega \right)
(\#eq:partialmaxloglik)\notag
\end{split}
\end{equation}

where $S = \sum_{i = 1}^{n}\left(X_{i} - \bar{X}\right)\left(X_{i} - \bar{X}\right)'/n$ is the usual sample estimator for the covariance matrix.

In addition to $\mu$, we could also find the maximum likelihood estimator for $\Omega$ and use that as our estimator. By setting the gradient of the partially maximized log-likelihood equal to zero, we find that

\begin{equation}
\begin{split}
\hat{\Omega}^{MLE} &= \arg\max_{\Omega \in S_{+}^{p}}\left\{ \frac{n}{2}\log\left| \Omega \right| - \frac{n}{2}tr\left(S\Omega \right) \right\} \\
  &= \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| \right\} \\
  &= S^{-1}
\end{split}
(\#eq:omegamle)\notag
\end{equation}

so that the MLE for $\Omega$, when it exists, is $\hat{\Omega}^{MLE} = S^{-1} = \left[\sum_{i = 1}^{n}\left(X_{i} - \bar{X}\right)\left(X_{i} - \bar{X}\right)'/n \right]$. The reality, however, is that this object does *not* always exist. In settings where the number of observations is exceeded by the number of features in a sample, the sample covariance matrix, $S$, is rank deficient and no longer invertible.

For this reason, many papers in the last decade have proposed shrinkage estimators of the precision matrix similar to the shrinkage estimators in regression settings. That is, instead of minimizing solely the negative log-likelihood function, authors have proposed minimizing the gaussian log-likelihood *plus* a penalty - often a function of the precision matrix. These methods effectively force the eigen values of the precision matrix, $\Omega$, to be strictly greater than zero.

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + P\left( \Omega \right) \right\}
(\#eq:omegapen)
\end{equation}

This penalty, which we will denote as $P\left(\Omega\right)$, is typically a variation of the ridge penalty $P\left(\Omega \right) = \lambda\|\Omega \|_{F}^{2}/2$ or the lasso penalty $P\left(\Omega \right) = \lambda\left\|\Omega\right\|_{1}$. The authors, @yuan2007model, were the ones who initially proposed precision matrix estimation by the lasso-penalized gaussian likelihood which they originally defined as

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\sum_{i \neq j}\left|\Omega_{ij}\right| \right\}
(\#eq:omegapenlasso)
\end{equation}

so as not to penalize the diagonal elements of the precision matrix estimate. Other papers published on the lasso-penalized gaussian likelihood precision matrix estimator include @rothman2008sparse and @friedman2008sparse and many efficient algorithms have been proposed to solve for $\hat{\Omega}$ in this setting. However, the most popular method is still the graphical lasso algorithm (sometimes referred to as glasso) introduced by @friedman2008sparse. Their method utilizes an iterative block-wise coordinate descent algorithm that builds upon the coordinate descent algorithm use in lasso regression.

Other papers published that propose methods with non-lasso penalizations like the Frobenius norms include @rothman2014existence, @witten2009covariance, and @price2015ridge. In fact, the latter two papers show that the estimation procedure can be solved in closed-form. Non-convex penalties were considered in @lam2009sparsistency and @fan2009network.

The penalty we will explore here and in the next section is one that is a convex combination of the lasso and the ridge penalties:

\begin{equation}
P\left( \Omega \right) = \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right]
(\#eq:penelastic)\notag
\end{equation}

where $0 \leq \alpha \leq 1$ is an additional tuning parameter (similar to $\lambda$). Clearly, when $\alpha = 0$ this penalty reduces to a ridge penalty and when $\alpha = 1$ it reduces to a lasso penalty. This penalty, know as the *elastic-net* penalty, was explored by Hui Zou and Trevor Hastie [@zou2005regularization] in the context of regression and is the penalty used in the popular penalized regression package `glmnet`. However, no published work has explored the elastic-net penalty in the context of precision matrix estimation. This penalty allows for increased flexibility and we will show how to solve the following optimization problem in the next section using the ADMM algorithm.

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in S_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega \right| + \lambda\left[\frac{1 - \alpha}{2}\left\| \Omega \right|_{F}^{2} + \alpha\left\| \Omega \right\|_{1} \right] \right\}
(\#eq:optimelastic)
\end{equation}

## ADMM Algorithm

ADMM stands for the alternating direction method of multipliers and this algorithm was largely popularized by Stephen Boyd and his fellow authors in the book *Distributed Optimization and Statistical Learning via the Alternating Direction Method of Multipliers* [@boyd2011distributed]. As the authors state in the text, the "ADMM is an algorithm that is intended to blend the decomposability of dual ascent with the superior convergence properties of the method of multipliers."

The ADMM algorithm - thanks to it's flexibility - is particularly well-suited to solve penalized-likelihood optimization problems that arise naturally in several statistics and machine learning applications.

In general, suppose we want to solve an optimization problem of the following form:

\begin{equation}
\begin{split}
  \mbox{minimize } f(x) + g(z) \\
  \mbox{subject to } Ax + Bz = c
\end{split}
(\#eq:boyd)\notag
\end{equation}

where $x \in \mathbb{R}^{n}, z \in \mathbb{R}^{m}, A \in \mathbb{R}^{p \times n}, B \in \mathbb{R}^{p \times m}$, $c \in \mathbb{R}^{p}$, and $f$ and $g$ are assumed to be convex functions (following @boyd2011distributed, the estimation procedure will be introduced in vector form though we could just as easily take $x$ and $z$ to be matrices). In addition to penalized precision matrix estimation, optimization problems like this arise naturally in several statistics and machine learning applications -- particularly regularization methods. For instance, we could take $f$ to be the squared error loss, $g$ to be the $l_{2}$-norm, $c$ to be equal to zero and $A$ and $B$ to be identity matrices to solve the ridge regression optimization problem. In all cases, our goal is to find $x^{*} \in \mathbb{R}^{n}$ and $z^{*} \in \mathbb{R}^{m}$ that achieves the infimum $p^{*}$:

\begin{equation}
p^{*} = inf\left\{ f(x) + g(z) | Ax + Bz = c \right\}
(\#eq:infimum)\notag
\end{equation}

To do so, the ADMM algorithm uses the *augmented lagrangian*

\begin{equation}
L_{\rho}(x, z, y) = f(x) + g(z) + y'(Ax + Bz - c) + \frac{\rho}{2}\left\| Ax + Bz - c \right\|_{2}^{2}
(\#eq:auglagrange)\notag
\end{equation}

where $y \in \mathbb{R}^{p}$ is the lagrange multiplier and $\rho > 0$ is a scalar. Clearly any minimizer, $p^{*}$, under the augmented lagrangian is equivalent to that of the lagrangian since any feasible point $(x, z)$ satisfies the constraint $\rho\left\| Ax + Bz - c \right\|_{2}^{2}/2 = 0$.

The algorithm consists of the following repeated iterations:

\begin{equation}
\begin{split}
  x^{k + 1} &= \arg\min_{x \in \mathbb{R}^{n}}L_{\rho}(x, z^{k}, y^{k}) \\
  z^{k + 1} &= \arg\min_{z \in \mathbb{R}^{m}}L_{\rho}(x^{k + 1}, z, y^{k}) \\
  y^{k + 1} &= y^{k} + \rho(Ax^{k + 1} + Bz^{k + 1} - c)
\end{split}  
(\#eq:admmalgo)\notag
\end{equation}

In the context of precision matrix estimation, we can let $f$ be equal to the non-penalized likelihood, $g$ be equal to $P\left( \Omega \right)$, and use the constraint $\Omega$ equal to some $Z$ so that the lagrangian is

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right]
(\#eq:lagrange2)\notag
\end{equation}

and the augmented lagrangian is

\begin{equation}
L_{\rho}(\Omega, Z, \Lambda) = f\left(\Omega\right) + g\left(Z\right) + tr\left[\Lambda\left(\Omega - Z\right)\right] + \frac{\rho}{2}\left\|\Omega - Z\right\|_{F}^{2}
(\#eq:auglagrange2)\notag
\end{equation}

The ADMM algorithm is now the following:

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  \Lambda^{k + 1} &= \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
\end{align}

### Algorithm

Set $k = 0$ and initialize $Z^{0}, \Lambda^{0}$, and $\rho$. Repeat steps 1-3 until convergence:

1. Decompose $S + \Lambda^{k} - \rho Z^{k} = VQV'$ via spectral decomposition^[Proof of \@ref(eq:omegaalgo) in section \@ref(proofomegaalgo).]. Then

\begin{equation}
\Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + \left( Q^{2} + 4\rho I_{p} \right)^{1/2} \right]V'
(\#eq:omegaalgo)
\end{equation}

2. Elementwise soft-thresholding for all $i = 1,..., p$ and $j = 1,..., p$^[Proof of \@ref(eq:ZZalgo) in section \@ref(proofZZalgo).].

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
(\#eq:ZZalgo)
\end{split}
\end{equation}

3. Update $\Lambda^{k + 1}$.

\begin{equation}
\Lambda^{k + 1} = \Lambda^{k} + \rho\left( \Omega^{k + 1} - Z^{k + 1} \right)
(\#eq:lamalgo)\notag
\end{equation}

where $\mbox{soft}(a, b) = \mbox{sign}(a)(\left| a \right| - b)_{+}$.

@witten2009covariance; @price2015ridge

### Scaled-Form ADMM

There is another popular, alternate form of the ADMM algorithm used by scaling the dual variable ($\Lambda^{k}$). Let us define $R^{k} = \Omega - Z^{k}$ and $U^{k} = \Lambda^{k}/\rho$.

\begin{equation}
\begin{split}
  tr\left[ \Lambda^{k}\left( \Omega - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} &= tr\left[ \Lambda^{k}R^{k} \right] + \frac{\rho}{2}\left\| R^{k} \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + \Lambda^{k}/\rho \right\|_{F}^{2} - \frac{\rho}{2}\left\| \Lambda^{k}/\rho \right\|_{F}^{2} \\
  &= \frac{\rho}{2}\left\| R^{k} + U^{k} \right\|_{F}^{2} - \frac{\rho}{2}\left\| U^{k} \right\|_{F}^{2}
(\#eq:scaled)\notag
\end{split}
\end{equation}

Therefore, a scaled-form can now be written as

\begin{align}
  \Omega^{k + 1} &= \arg\min_{\Omega \in \mathbb{R}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\rho}{2}\left\| \Omega - Z^{k} + U^{k} \right\|_{F}^{2} \right\} \\
  Z^{k + 1} &= \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z + U^{k} \right\|_{F}^{2} \right\} \\
  U^{k + 1} &= U^{k} + \Omega^{k + 1} - Z^{k + 1}
\end{align}

And more generally (in vector form),

\begin{align}
  x^{k + 1} &= \arg\min_{x}\left\{ f(x) + \frac{\rho}{2}\left\| Ax + Bz^{k} - c + u^{k} \right\|_{2}^{2} \right\} \\
  z^{k + 1} &= \arg\min_{z}\left\{ g(z) + \frac{\rho}{2}\left\| Ax^{k + 1} + Bz - c + u^{k} \right\|_{2}^{2} \right\} \\
  u^{k + 1} &= u^{k} + Ax^{k + 1} + Bz^{k + 1} - c
\end{align}

Note that there are limitations to using this method. Because the dual variable is scaled by $\rho$ (the step size), this form limits one to using a constant step size without making further adjustments to $U^{k}$. It has been shown in the literature that a dynamic step size (like the one used in `ADMMsigma`) can significantly reduce the number of iterations required for convergence.

### Stopping Criterion

In discussing the optimality conditions and stopping criterion, we will follow the steps outlined in @boyd2011distributed and cater them to precision matrix estimation.

Below we have three optimality conditions:

1. Primal:

\begin{equation}
\Omega^{k + 1} - Z^{k + 1} = 0
(\#eq:admmprimal)\notag
\end{equation}

2. Dual:

\begin{equation}
\begin{split}
  0 &\in \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} \\
  0 &\in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1}
(\#eq:admmdual)\notag
\end{split}
\end{equation}

The first dual optimality condition is a result of taking the sub-differential of the lagrangian (non-augmented) with respect to $\Omega^{k + 1}$ (note that we must honor the symmetric constraint of $\Omega^{k + 1}$) and the second is a result of taking the sub-differential of the lagrangian with respect to $Z^{k + 1}$ (no symmetric constraint).

We will define the left-hand side of the first condition as the primal residual $r^{k + 1} = \Omega^{k + 1} - Z^{k + 1}$. At convergence, optimality conditions require that $r^{k + 1} \approx 0$. The second residual we will define is the dual residual $s^{k + 1} = \rho\left( Z^{k + 1} - Z^{k} \right)$. This residual is derived from the following:

Because $\Omega^{k + 1}$ is the minimizer of the augmented lagragian^[Proof of \@ref(eq:dualresidual) in section \@ref(proofdualresidual).], this implies that $0 \in \partial L_{p}\left( \Omega, Z^{k}, \Lambda^{k} \right)$ which further implies that

\begin{equation}
0 \in \rho\left( Z^{k + 1} - Z^{k} \right)
(\#eq:dualresidual)
\end{equation}

Like the primal residual, at convergence optimality conditions require that $s^{k + 1} \approx 0$. Note that the second dual optimality condition is always satisfied^[Proof of \@ref(eq:dualopt) in section \@ref(proofdualopt).]:

\begin{equation}
\begin{split}
  0 \in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
(\#eq:dualopt)
\end{split}
\end{equation}

One possible stopping criterion is to set $\epsilon^{rel} = \epsilon^{abs} = 10^{-3}$ and stop the algorithm when $\epsilon^{pri} \leq \left\| r^{k + 1} \right\|_{F}$ and $\epsilon^{dual} \leq \left\| s^{k + 1} \right\|_{F}$ where

\begin{equation}
\begin{split}
  \epsilon^{pri} &= p\epsilon^{abs} + \epsilon^{rel}\max\left\{ \left\| \Omega^{k + 1} \right\|_{F}, \left\| Z^{k + 1} \right\|_{F} \right\} \\
  \epsilon^{dual} &= p\epsilon^{abs} + \epsilon^{rel}\left\| \Lambda^{k + 1} \right\|_{F}
(\#eq:admmstopping)\notag
\end{split}
\end{equation}

## Simulations

In the simulations below, we generate data from a number of different oracle precision matrices with various structures. For each data-generating procedure, the `ADMMsigma()` function was run using 5-fold cross validation. After 20 replications, the cross validation errors were totalled and the optimal tuning parameters were selected (results below the top figure). These results are compared with the Kullback Leibler (KL) losses between the estimates and the oracle precision matrix (bottom figure).

We can see below that our cross validation procedure consistently chooses tuning parameters that are close to the optimal parameters with respsect to the oracle.

HOW DID WE GENERATE THE DATA?

### Plots

#### Compound Symmetric: P = 100, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE}

# oracle precision matrix
Omega = matrix(0.9, ncol = 100, nrow = 100)
diag(Omega = 1)

# generate covariance matrix
S = qr.solve(Omega)

# generate data
Z = matrix(rnorm(100*50), nrow = 50, ncol = 100)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/compound_N50_P100.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim1, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/compound_N50_P100.png")

```

#### Compound Symmetric: P = 10, N = 1000

```{r, message = FALSE, echo = TRUE, eval = FALSE}

# oracle precision matrix
Omega = matrix(0.9, ncol = 10, nrow = 10)
diag(Omega = 1)

# generate covariance matrix
S = qr.solve(Omega)

# generate data
Z = matrix(rnorm(10*1000), nrow = 1000, ncol = 10)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/compound_N1000_P10.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim2, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/compound_N1000_P10.png")

```

<!-- #### Dense: P = 100, N = 50 -->

<!-- ```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE} -->

<!-- # generate eigen values -->
<!-- eigen = c(rep(1000, 5, rep(1, 100 - 5))) -->

<!-- # randomly generate orthogonal basis (via QR) -->
<!-- Q = matrix(rnorm(100*100), nrow = 100, ncol = 100) %>% qr %>% qr.Q -->

<!-- # generate covariance matrix -->
<!-- S = Q %*% diag(eigen) %*% t(Q) -->

<!-- # generate data -->
<!-- Z = matrix(rnorm(100*50), nrow = 50, ncol = 100) -->
<!-- out = eigen(S, symmetric = TRUE) -->
<!-- S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors) -->
<!-- X = Z %*% S.sqrt -->

<!-- ``` -->
<!-- <br>\vspace{0.5cm} -->

<!-- <!-- <center> --> -->
<!-- <!-- ![Figure caption here.](images/repsKLdenseQR_N50_P100.png){ --> -->
<!-- <!-- width=90% } --> -->
<!-- <!-- </center> --> -->
<!-- <!-- } --> -->

<!-- ```{r admmsim3, eval = T, echo = F, fig.cap = "Figure caption here."} -->

<!-- knitr::include_graphics("images/repsKLdenseQR_N50_P100.png") -->

<!-- ``` -->

<!-- #### Dense: P = 10, N = 50 -->

<!-- ```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE} -->

<!-- # generate eigen values -->
<!-- eigen = c(rep(1000, 5, rep(1, 10 - 5))) -->

<!-- # randomly generate orthogonal basis (via QR) -->
<!-- Q = matrix(rnorm(10*10), nrow = 10, ncol = 10) %>% qr %>% qr.Q -->

<!-- # generate covariance matrix -->
<!-- S = Q %*% diag(eigen) %*% t(Q) -->

<!-- # generate data -->
<!-- Z = matrix(rnorm(10*50), nrow = 50, ncol = 10) -->
<!-- out = eigen(S, symmetric = TRUE) -->
<!-- S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors) -->
<!-- X = Z %*% S.sqrt -->

<!-- ``` -->
<!-- <br>\vspace{0.5cm} -->

<!-- <!-- <center> --> -->
<!-- <!-- ![Figure caption here.](images/repsKLdense_N50_P10.png){ --> -->
<!-- <!-- width=90% } --> -->
<!-- <!-- </center> --> -->

<!-- ```{r admmsim4, eval = T, echo = F, fig.cap = "Figure caption here."} -->

<!-- knitr::include_graphics("images/repsKLdense_N50_P10.png") -->

<!-- ``` -->

#### Tridiagonal: P = 100, N = 50

```{r, message = FALSE, echo = TRUE, eval = FALSE, tidy = FALSE}

# generate covariance matrix
# (can confirm inverse is tri-diagonal)
S = matrix(0, nrow = 100, ncol = 100)
for (i in 1:100){
  for (j in 1:100){
    S[i, j] = 0.7^abs(i - j)
  }
}

# generate data
Z = matrix(rnorm(10*50), nrow = 50, ncol = 10)
out = eigen(S, symmetric = TRUE)
S.sqrt = out$vectors %*% diag(out$values^0.5) %*% t(out$vectors)
X = Z %*% S.sqrt

```
<br>\vspace{0.5cm}

<!-- <center> -->
<!-- ![Figure caption here.](images/repsKLtridiag_N50_P100.png){ -->
<!-- width=90% } -->
<!-- </center> -->

```{r admmsim5, eval = T, echo = F, fig.cap = "Figure caption here."}

knitr::include_graphics("images/repsKLtridiag_N50_P100.png")

```