### 2019-04-20 05-appendix.Rmd

# (APPENDIX) Appendix {-}

# Appendix

### Proof of \@ref(eq:omegaalgo) {#proofomegaalgo}

@witten2009covariance and @price2015ridge showed that the following optimization problem can be solved in closed-form. We outline their steps here.

\begin{equation}
\Omega^{k + 1} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\}
(\#eq:appendixomegaloglik)\notag
\end{equation}

First start by taking the gradient with respect to $\Omega$:

\begin{equation}
\begin{split}
  \nabla_{\Omega}&\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[\Lambda^{k}\left(\Omega - Z^{k}\right)\right] + \frac{\rho}{2}\left\| \Omega - Z^{k} \right\|_{F}^{2} \right\} \\
  &= S - \Omega^{-1} + \Lambda^{k} + \rho\left( \Omega - Z^{k} \right)
(\#eq:appendixomegagradient)\notag
\end{split}
\end{equation}

Note that because all of the variables are symmetric, we can ignore the symmetric constraint when deriving the gradient. Next set the gradient equal to zero and decompose $\Omega^{k + 1} = VDV'$ using spectral decomposition so that $D$ is a diagonal matrix with diagonal elements equal to the eigen values of $\Omega^{k + 1}$ and $V$ is the matrix with corresponding eigen vectors as columns so that

\begin{equation}
S + \Lambda^{k} - \rho Z^{k} = (\Omega^{k + 1})^{-1} - \rho \Omega^{k + 1} = VD^{-1}V' - \rho VDV' =  V\left(D^{-1} - \rho D\right)V'\notag
\end{equation}

This equivalence implies that

\begin{equation}
\phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right) = \frac{1}{\phi_{j}(\Omega^{k + 1})} - \rho\phi_{j}(\Omega^{k + 1})\notag
\end{equation}

where $\phi_{j}(\cdot)$ is the $j$th eigen value.

\begin{equation}
\begin{split}
  &\Rightarrow \rho\phi_{j}^{2}(\Omega^{k + 1}) + \phi_{j}\left( S + \Lambda^{k} - \rho Z^{k} \right)\phi_{j}(\Omega^{k + 1}) - 1 = 0 \\
  &\Rightarrow \phi_{j}(\Omega^{k + 1}) = \frac{-\phi_{j}(S + \Lambda^{k} - \rho Z^{k}) \pm \sqrt{\phi_{j}^{2}(S + \Lambda^{k} - \rho Z^{k}) + 4\rho}}{2\rho}
\end{split}
\notag
\end{equation}

In summary, if we decompose $S + \Lambda^{k} - \rho Z^{k} = VQV'$ then

\begin{equation}
\Omega^{k + 1} = \frac{1}{2\rho}V\left[ -Q + (Q^{2} + 4\rho I_{p})^{1/2}\right] V'
(\#eq:appendixomegaproof)
\end{equation}

### Proof of \@ref(eq:ZZalgo) {#proofZZalgo}

Solve the optimization problem

\begin{equation}
Z^{k + 1} = \arg\min_{Z \in \mathbb{S}^{p}}\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\}
(\#eq:appendixZZloglik)\notag
\end{equation}

Start by taking the subgradient with respect to $Z$:

\begin{equation}
\begin{split}
  \partial&\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] + tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \partial\left\{ \lambda\left[ \frac{1 - \alpha}{2}\left\| Z \right\|_{F}^{2} + \alpha\left\| Z \right\|_{1} \right] \right\} + \nabla_{\Omega}\left\{ tr\left[\Lambda^{k}\left(\Omega^{k + 1} - Z\right)\right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z \right\|_{F}^{2} \right\} \\
  &= \lambda(1 - \alpha)Z + \mbox{sign}(Z)\lambda\alpha - \Lambda^{k} - \rho\left( \Omega^{k + 1} - Z \right)
(\#eq:appendixZZgradient)\notag
\end{split}
\end{equation}

where sign is the elementwise sign operator. By setting the gradient/sub-differential equal to zero, we have that

\begin{equation}
Z_{ij}^{k + 1} = \frac{1}{\lambda(1 - \alpha) + \rho}\left( \rho \Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} - \mbox{sign}(Z_{ij}^{k + 1})\lambda\alpha \right)
(\#eq:ZZsoftproof)\notag
\end{equation}

for all $i = 1,..., p$ and $j = 1,..., p$. We observe two scenarios:

- If $Z_{ij}^{k + 1} > 0$ then

\begin{equation}
\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} > \lambda\alpha \notag
\end{equation}

- If $Z_{ij}^{k + 1} < 0$ then

\begin{equation}
\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} < -\lambda\alpha \notag
\end{equation}

This implies that $\mbox{sign}(Z_{ij}) = \mbox{sign}(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k})$. Putting all the pieces together, we arrive at

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{sign}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}\right)\left( \left| \rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k} \right| - \lambda\alpha \right)_{+} \\
&= \frac{1}{\lambda(1 - \alpha) + \rho}\mbox{soft}\left(\rho\Omega_{ij}^{k + 1} + \Lambda_{ij}^{k}, \lambda\alpha\right)
(\#eq:appendixZZproof)\notag
\end{split}
\end{equation}

where soft is the soft-thresholding function.

### Proof of the dual residual {#proofdualresidual}

Here we want to show that $0 \in \rho\left(Z^{k + 1} - Z^{k}\right)$ which is a direct result of the fact that $\Omega^{k + 1}$ is the minimizer of the augmented lagrangian and $0 \in f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1}$.

\begin{equation}
\begin{split}
  0 &\in \partial \left\{ f\left(\Omega^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k} \right) \right] + \frac{\rho}{2}\left\| \Omega^{k + 1} - Z^{k} \right\|_{F}^{2} \right\} \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} + Z^{k + 1} - Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k} + \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  &= \partial f\left(\Omega^{k + 1}\right) + \Lambda^{k + 1} + \rho\left(Z^{k + 1} - Z^{k}\right) \\
  \Rightarrow 0 &\in \rho\left( Z^{k + 1} - Z^{k} \right)
(\#eq:appendixdualresidual)\notag
\end{split}
\end{equation}

### Proof of second dual optimality condition {#proofdualopt}

Here we use the primal optimality condition $\Omega^{k + 1} - Z^{k + 1} = 0$ to show that the second dual optimality condition $0 \in \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1}$ is always satisfied.

\begin{equation}
\begin{split}
  0 &\in \partial \left\{ g\left(Z^{k + 1}\right) + tr\left[ \Lambda^{k}\left( \Omega^{k + 1} - Z^{k + 1} \right) \right] + \rho\left\| \Omega^{k + 1} - Z^{k + 1} \right\|_{F}^{2} \right\} \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k} - \rho\left(\Omega^{k + 1} - Z^{k + 1}\right) \\
  &= \partial g\left(Z^{k + 1}\right) - \Lambda^{k + 1} \\
(\#eq:appendixdualopt)\notag
\end{split}
\end{equation}

### Explanation of majorizing function \@ref(eq:approx) {#taylorsexp}

To see why this particular function was used, consider the Taylor's expansion of $\rho\left\|A\Omega B - Z^{k} - C\right\|_{F}^{2}/2$:

\begin{equation}
\begin{split}
  \frac{\rho}{2}\left\| A\Omega B - Z^{k} - C \right\|_{F}^{2} &\approx \frac{\rho}{2}\left\| A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2} \\
  &+ \frac{\rho}{2}vec\left( \Omega - \Omega^{k}\right)'\left(A'A \otimes BB'\right)vec\left(\Omega - \Omega^{k}\right) \\
  &+ \rho vec\left(\Omega - \Omega^{k}\right)'vec\left(BB'\Omega^{k}A'A - B(Z^{k})'A - BC'A \right)
(\#eq:taylors)\notag
\end{split}
\end{equation}

Note that the gradient and hessian, respectively, are

\begin{equation}
  \nabla_{\Omega}\left\{ \frac{\rho}{2}\left\|A\Omega B - Z - C\right\|_{F}^{2} \right\} = \rho BB'\Omega A'A - \rho BZ'A - \rho BC'A
(\#eq:gradient)\notag
\end{equation}

\begin{equation}
\nabla_{\Omega}^{2}\left\{ \frac{\rho}{2}\left\|A\Omega B - Z - C \right\|_{F}^{2} \right\} = \rho\left(A'A \otimes BB' \right)
(\#eq:hessian)\notag
\end{equation}

This implies that

\begin{equation}
\begin{split}
  \frac{\rho}{2}\left\| A\Omega B - Z^{k} - C \right\|_{F}^{2} &+ \frac{\rho}{2}vec\left(\Omega - \Omega^{k} \right)'Q\left(\Omega - \Omega^{k} \right) \\
  &\approx \frac{\rho}{2}\left\| A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2} + \frac{\rho}{2}vec\left(\Omega - \Omega^{k} \right)'Q\left(\Omega - \Omega^{k} \right) \\
  &+ \frac{\rho}{2}vec\left( \Omega - \Omega^{k}\right)'\left(A'A \otimes BB'\right)vec\left(\Omega - \Omega^{k}\right) \\
  &+ \rho vec\left(\Omega - \Omega^{k}\right)' vec\left(BB'\Omega^{k}A'A - B(Z^{k})'A - BC'A \right) \\
  &= \frac{\rho}{2}\left\| A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2} + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k}\right\|_{F}^{2} \\
  &+ \rho tr\left[\left(\Omega - \Omega^{k}\right)\left(BB'\Omega^{k}A'A - B(Z^{k})'A - BC'A \right)\right]
(\#eq:penapprox)\notag
\end{split}
\end{equation}

Let us now plug in this equality into our optimization problem that includes the augmented lagrangian:

\begin{equation}
\begin{split}
  \hat{\Omega}^{k + 1} &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\tilde{L}_{\rho}(\Omega, Z^{k}, \Lambda^{k}) \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})'(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} vec\left(\Omega - \Omega^{k}\right)'\rho Q\left(\Omega - \Omega^{k}\right)/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left(S\Omega\right) - \log\left|\Omega\right| + tr\left[(\Lambda^{k})'(A\Omega B - Z^{k} - C) \right] + \rho\left\|A\Omega^{k} B - Z^{k} - C \right\|_{F}^{2}/2 \end{matrix}\right. \\
  &+ \left.\begin{matrix} \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 + tr\left[\rho\left(\Omega - \Omega^{k}\right)\left(BB'\Omega^{k}A'A - B(Z^{k})'A - BC'A \right)\right] \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{\begin{matrix}
 tr\left[\left(S + \rho A'(A\Omega^{k}B - Z^{k} - C + \Lambda^{k}/\rho)B' \right)\Omega\right] \end{matrix}\right. \\
  &- \left.\begin{matrix} \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \end{matrix}\right\} \\
  &= \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{
 tr\left[\left(S + G^{k} \right)\Omega\right] - \log\left|\Omega\right| + \rho\tau\left\|\Omega - \Omega^{k}\right\|_{F}^{2}/2 \right\} \\
(\#eq:omegaapprox)\notag
\end{split}
\end{equation}

where $G^{k} = \rho A'(A\Omega^{k}B - Z^{k} - C + \Lambda^{k}/\rho)B'$.

### Proof of \@ref(eq:Omegak) {#proofOmegak}

We show that the following optimization problem can be solved in closed-form similar to \@ref(eq:appendixomegaproof).

\begin{equation}
\Omega^{k + 1} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{tr\left[\left(S + G^{k}\right)\Omega\right] - \log\left|\Omega\right| + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k}\right\|_{F}^{2} \right\}
(\#eq:appendixomegaloglik2)\notag
\end{equation}

First start by taking the gradient with respect to $\Omega$:

\begin{equation}
\begin{split}
  &\nabla_{\Omega}\left\{tr\left[\left(S + G^{k}\right)\Omega\right] - \log\left|\Omega\right| + \frac{\rho\tau}{2}\left\|\Omega - \Omega^{k}\right\|_{F}^{2} \right\} \\
  &= 2S - S\circ I_{p} + G^{k} + (G^{k})' - G^{k}\circ I_{p} - 2\Omega^{-1} + \Omega^{-1}\circ I_{p} \\
  &+ \frac{\rho\tau}{2}\left[2\Omega - 2(\Omega^{k})' + 2\Omega' - 2\Omega^{k} - 2(\Omega - \Omega^{k})'\circ I_{p} \right]
(\#eq:appendixomegagradient2)\notag
\end{split}
\end{equation}

Note that we need to honor the symmetric constraint given by $\Omega$. By setting the gradient equal to zero and multiplying all off-diagonal elements by $1/2$, this simplifies to

\begin{equation}
S + \frac{1}{2}\left(G^{k} + (G^{k})'\right) - \rho\tau\Omega^{k} = (\Omega^{k + 1})^{-1} - \rho\tau\Omega^{k + 1} \notag
\end{equation}

We can then decompose using spectral decomposition $\Omega^{k + 1} = VDV'$ where $D$ is a diagonal matrix with diagonal elements equal to the eigen values of $\Omega^{k + 1}$ and $V$ is the matrix with corresponding eigen vectors as columns.

\begin{equation}
S + \frac{1}{2}\left(G^{k} + (G^{k})'\right) - \rho\tau\Omega^{k} = VD^{-1}V' - \rho\tau VDV' = V\left(D^{-1} - \rho\tau D\right)V' \notag
\end{equation}

This equivalence implies that

\begin{equation}
\phi_{j}\left( D^{k} \right) = \frac{1}{\phi_{j}(\Omega^{k + 1})} - \rho\tau\phi_{j}(\Omega^{k + 1}) \notag
\end{equation}

where $\phi_{j}(\cdot)$ is the $j$th eigen value and $D^{k} = S + \left(G^{k} + (G^{k})'\right)/2 - \rho\tau\Omega^{k}$. Therefore

\begin{equation}
\begin{split}
  &\Rightarrow \rho\tau\phi_{j}^{2}(\Omega^{k + 1}) + \phi_{j}\left( D^{k} \right)\phi_{j}(\Omega^{k + 1}) - 1 = 0 \\
  &\Rightarrow \phi_{j}(\Omega^{k + 1}) = \frac{-\phi_{j}(D^{k}) \pm \sqrt{\phi_{j}^{2}(D^{k}) + 4\rho\tau}}{2\rho\tau}
\end{split}
\notag
\end{equation}

In summary, if we decompose $S + \left(G^{k} + (G^{k})'\right)/2 - \rho\tau\Omega^{k} = VQV'$ then

\begin{equation}
\Omega^{k + 1} = \frac{1}{2\rho\tau}V\left[ -Q + (Q^{2} + 4\rho\tau I_{p})^{1/2}\right] V'
(\#eq:appendixomegaproof2)\notag
\end{equation}

### Proof of \@ref(eq:Zk) {#proofZk}

Solve the optimization problem

\begin{equation}
Z^{k + 1} = \arg\min_{Z \in \mathbb{R}^{n \times r}}\left\{ \lambda\left\| Z \right\|_{1} + tr\left[(\Lambda^{k})'\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\}
(\#eq:appendixZZloglik2)\notag
\end{equation}

Start by taking the subgradient with respect to $Z$:

\begin{equation}
\begin{split}
  \partial&\left\{ \lambda\left\| Z \right\|_{1} + tr\left[(\Lambda^{k})'\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\} \\
  &= \partial\left\{ \lambda\left\| Z \right\|_{1} \right\} + \nabla_{\Omega}\left\{ tr\left[(\Lambda^{k})'\left(A\Omega^{k + 1}B - Z - C\right)\right] + \frac{\rho}{2}\left\| A\Omega^{k + 1}B - Z - C \right\|_{F}^{2} \right\} \\
  &= \mbox{sign}(Z)\lambda - \Lambda^{k} - \rho\left( A\Omega^{k + 1}B - Z - C \right)
(\#eq:appendixZZgradient2)\notag
\end{split}
\end{equation}

where $\mbox{sign(Z)}$ is the elementwise sign operator. By setting the gradient/sub-differential equal to zero, we arrive at the following equivalence:

\begin{equation}
Z_{ij}^{k + 1} = \frac{1}{\rho}\left( \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} - Sign\left(Z_{ij}^{k + 1}\right)\lambda \right)
(\#eq:ZZsoftproof2)\notag
\end{equation}

for all $i = 1,..., p$ and $j = 1,..., p$. We observe two scenarios:

- If $Z_{ij}^{k + 1} > 0$ then

\begin{equation}
\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} > \lambda\alpha \notag
\end{equation}

- If $Z_{ij}^{k + 1} < 0$ then

\begin{equation}
\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} < -\lambda\alpha \notag
\end{equation}

This implies that $\mbox{sign}\left(Z_{ij}^{k + 1}\right) = \mbox{sign}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}\right)$. Putting all the pieces together, we arrive at

\begin{equation}
\begin{split}
Z_{ij}^{k + 1} &= \frac{1}{\rho}\mbox{sign}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}\right)\left( \left| \rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k} \right| - \lambda \right)_{+} \\
&= \frac{1}{\rho}\mbox{soft}\left(\rho\left(A\Omega_{ij}^{k + 1}B - C\right) + \Lambda_{ij}^{k}, \lambda\right)
(\#eq:appendixZZproof2)\notag
\end{split}
\end{equation}

where soft is the soft-thresholding function.

### Proof of \@ref(eq:stopproof) {#proofstopproof}

Here we want to show that $0 \in \rho\left( B(Z^{k + 1} - Z^{k})'A + A'(Z^{k + 1} - Z^{k})B' \right)/2$ which is a direct result of the fact that $0 \in \partial f\left(\Omega^{k + 1} \right) + \left(B(\Lambda^{k + 1})'A + A'\Lambda^{k + 1}B' \right)/2$.

\begin{equation}
\begin{split}
  0 &\in \partial \left\{ f\left(\Omega^{k + 1}\right) + tr\left[ \Lambda^{k}\left( A\Omega^{k + 1}B - Z^{k} - C \right) \right] + \rho\left\| A\Omega^{k + 1}B - Z^{k} - C \right\|_{F}^{2}/2 \right\} \\
  &= \partial f\left(\Omega^{k + 1} \right) + \left(B(\Lambda^{k})'A + A'\Lambda^{k}B' \right)/2 + \rho\left( BB'\Omega^{k + 1}A'A + A'A\Omega^{k + 1}BB' \right)/2 \\
  &- \rho\left( A'(Z^{k} + C)B' + B(Z^{k} + C)'A \right)/2 \\
  &= \partial f\left(\Omega^{k + 1} \right) + \left(B(\Lambda^{k})'A + A'\Lambda^{k}B' \right)/2 \\
  &+ \rho\left( B(B'\Omega^{k + 1}A' - (Z^{k})' - C')A + A'(A\Omega^{k + 1}B - Z^{k} - C)B' \right)/2 \\
  &= \partial f\left(\Omega^{k + 1} \right) + \left( B(\Lambda^{k})'A + A'\Lambda^{k}B' \right)/2 + \rho\left(A'(A\Omega^{k + 1}B - Z^{k + 1} + Z^{k + 1} - Z^{k} - C)B' \right)/2 \\
  &+ \rho\left(B(B'\Omega^{k + 1}A' - (Z^{k + 1})' + (Z^{k + 1})' - (Z^{k})' - C')A \right)/2 \\
  &= \partial f\left(\Omega^{k + 1} \right) + \left[ B\left((\Lambda^{k})' + \rho(B'\Omega^{k + 1}A' - (Z^{k + 1})' - C') \right)A \right]/2 \\
  &+ \left[ A'\left(\Lambda^{k} + \rho(A\Omega^{k + 1}B - Z^{k + 1} - c)B \right)B' \right]/2 + \rho\left(B(Z^{k + 1} - Z^{k})'A + A'(Z^{k + 1} - Z^{k})B' \right)/2 \\
  &= \partial f\left(\Omega^{k + 1} \right) + \left(B(\Lambda^{k + 1})'A + A'\Lambda^{k + 1}B' \right)/2 + \rho\left(B(Z^{k + 1} - Z^{k})'A + A'(Z^{k + 1} - Z^{k})B' \right)/2 \\
  \Rightarrow 0 &\in \rho\left( B(Z^{k + 1} - Z^{k})'A + A'(Z^{k + 1} - Z^{k})B' \right)/2
(\#eq:appendixstopproof)\notag
\end{split}
\end{equation}

### Proof of no closed-form solution {#proofpenregression}

Here we want to show that the following optimization problem cannot be solved in closed-form.

\begin{equation}
\hat{\Omega} = \arg\min_{\Omega \in \mathbb{S}_{+}^{p}}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\lambda}{2}\left\|\mathbb{X}\Omega\Sigma_{xy} - \mathbb{Y}\right\|_{F}^{2} \right\}
(\#eq:penregression2)\notag
\end{equation}

First start by taking the gradient with respect to $\Omega$:

\begin{equation}
\begin{split}
  \nabla_{\Omega}&\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\lambda}{2}\left\|\mathbb{X}\Omega\Sigma_{xy} - \mathbb{Y}\right\|_{F}^{2} \right\} \\
  &= \nabla_{\Omega}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\lambda}{2}tr\left[\left(\mathbb{X}\Omega\Sigma_{xy} - \mathbb{Y}\right)'\left(\mathbb{X}\Omega\Sigma_{xy} - \mathbb{Y}\right) \right] \right\} \\
  &= \nabla_{\Omega}\left\{ tr\left(S\Omega\right) - \log\left|\Omega\right| + \frac{\lambda}{2}tr\left(\Omega\mathbb{X}'\mathbb{X}\Omega\Sigma_{xy}\Sigma_{xy}' - 2\Omega\Sigma_{xy}\mathbb{Y}'\mathbb{X} \right) \right\} \\
  &= 2S - S\circ I_{p} - 2\Omega^{-1} + \Omega^{-1}\circ I_{p} + \lambda\mathbb{X}'\mathbb{X}\Omega\Sigma_{xy}\Sigma_{xy}' + \lambda\Sigma_{xy}\Sigma_{xy}'\Omega\mathbb{X}'\mathbb{X} \\
  &- \lambda\mathbb{X}'\mathbb{X}\Omega\Sigma_{xy}\Sigma_{xy}'\circ I_{p} - \Sigma_{xy}\mathbb{Y}'\mathbb{X} - \mathbb{X}'\mathbb{Y}\Sigma_{xy} + \Sigma_{xy}\mathbb{Y}'\mathbb{X}\circ I_{P} \\
  \Rightarrow 0 &= S - \hat{\Omega}^{-1} + \frac{\lambda}{2}\left(\mathbb{X}'\mathbb{X}\hat{\Omega}\Sigma_{xy}\Sigma_{xy}' + \Sigma_{xy}\Sigma_{xy}'\hat{\Omega}\mathbb{X}'\mathbb{X} \right) - \lambda\Sigma_{xy}\mathbb{Y}'\mathbb{X}
(\#eq:penregressionproof)\notag
\end{split}
\end{equation}

Note that we cannot isolate $\hat{\Omega}$.